### 2023-10-22 outputs/paper/paper.qmd
Update the paper with the introduction
First Draft

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

On May 23rd, the WHO chief Tedros announced that the COVID-19 ends and COVID-19 is no longer regarded as a global threat. At this point, the epidemic that lasted for three years has officially ended. During the three years, nearly 7.6 billion people got infected and about 6.9 million people lost their lives. With that said, the US seems to be one of the countries that had influenced by COVID-19 most, there was about 1.1 million deaths and 104 million infections. The mortality rate of COVID-19 in US is about 341 per 100,000, which is significantly higher than other western developed countries. Meanwhile, the COVID was entirely occured in the Biden's term and next year will be the Federal Election. It is worth to know how the COVID deaths is related to the politics and how to recover it, especially that we are now at the post-pandemic period.

In this paper, I aim to find the variations of death rate across different counties, to see how the counties with different socio-economic factors were affected by COVID differently. In addition, I will also include the political factors, to see whether the party that counties voted for will display any differences on COVID death.  There are three primary data sources which are American Community Survey collecting the socio-economic information for each county, the John-Hopkins public data for COVID cases and deaths for each county and the 2020 US Federal Election from Fox News. With capturing this variations, it can help us to access the inequality of mortality caused by COVID across different counties. Besides, people living at each county will also know whether their political reference will affect the death rate, this can significantly judge them about how they will vote on the 2024 Federal Election.

There will be four main parts in this paper: Data, Models, Results, Discussion and Conclusion. In the Data part, I will introduce the data used in this paper and high light the key variables. The Models and Results session will introduce the model that will be used in this paper and the results from it. Moreover, the results will be interpreted and provide the insights about COVID deaths rate. Lastly, I will conclude the entire paper and discussion the limitations and drawbacks.

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  
\[
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
\]
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\newpage

# References
https://news.un.org/en/story/2023/05/1136367
https://www.aljazeera.com/news/2023/5/11/three-years-1-1-million-deaths-covid-emergency-ending-in-us

### 2023-10-23 outputs/paper/paper.qmd
Update the introduction and data part

@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+In this paper, there are three main sources of data, which are American Community Survey (ACS), The Center for Systems Science and Engineering (CSSE) at John-Hopkins University, Fox News.
+These three sources contains different dimensional data in predicting the death rate of COVID in each county.
+
+## American Community Survey (ACS)
+
+The data from ACS contains the socio-economic information for each county.
+While there are many tables from ACS, I will only take three of them which are DP02, DP03 and DP05.
+All the tables are the 2021 five-year ACS estimates
+
+For DP02, it contains the data regarding the social characteristics.
+In this paper, I will only take the information regarding the educational attainment such as the percentage of residents in each county that have at a least bachelor degree.
+Compared to DP02, DP03 contains the information about the economic characteristics.
+To understand the deaths rate in terms of economy, I will grab the information about the proportion of residents with a private insurance, the mean household income and unemployment rate at each county.
+Finally, DP05 contains the demographic data and housing estimates.
+In this paper, I will take the total population, proportions of children and people above 85 and percentage of white and black residents.
+
+## CSSE at JHU
+The CSSE at JHU collects the worldwide COVID data since 2020 to 2023.
+All the raw data are available on their CSSE GitHub which can easily access.
+In particular, the COVID daily report since 2020 for each county on US are listed.
+However, they stop to update the data after March 9th, 2023 as no significant changes.
+The COVID data for US used in the paper is data collected on March 9th.
+
+In the data on March 9th, it contains the number of confirmed cases and deaths for each county in US, as well as the incident rate.
+To perform the data analysis later in this paper, I will only use the COVID cases and deaths in each county.
+
+## Fox News
+It contains the county-level election results during the 2020 Federal Election. For each county, it contains the number of votes for the demographic and republican party
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  
-\[
-\begin{align} 
+Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.\
+\[ \begin{align} 
 y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
@@
 \sigma &\sim \mbox{Exponential}(1)
-\end{align}
-\]
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+\end{align} \] We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
 
@@
 # References
-https://news.un.org/en/story/2023/05/1136367
-https://www.aljazeera.com/news/2023/5/11/three-years-1-1-million-deaths-covid-emergency-ending-in-us
+
+https://news.un.org/en/story/2023/05/1136367 https://www.aljazeera.com/news/2023/5/11/three-years-1-1-million-deaths-covid-emergency-ending-in-us

### 2023-10-24 outputs/paper/paper.qmd
Update the data part

@@
 ## Fox News
-It contains the county-level election results during the 2020 Federal Election. For each county, it contains the number of votes for the demographic and republican party
+
+It contains the county-level election results during the 2020 Federal Election. For each county, it contains the number of votes for the demographic and republican party.
+
+## Mergin the three sources
+
+To predict the COVID death rate across counties with various socio-economic features, I will merge the the five data sets from the above three data sources by the county and state.
+In the merged data set, each row represent a US county and provides the number of COVID cases and deaths.
+In addition, it also shows the demographic information such as the total population and voting pattern such as the number of votes for democratic and republican party.
+To better understand the graph, I will several graphs.
 

### 2023-10-28 scripts/01-download_data.R
Update 01-download_data.R
Download the api change document

@@ -18,6 +18,8 @@ url03 <- "https://api.census.gov/data/2021/acs/acs5/profile?get=group(DP03)&for=
 DP03 <- read.csv(url(url03), stringsAsFactors = FALSE)
 url05 <- "https://api.census.gov/data/2021/acs/acs5/profile?get=group(DP05)&for=county:*"
 DP05 <- read.csv(url(url05), stringsAsFactors = FALSE)
+url06 <- "https://www2.census.gov/data/api-documentation/2021-5yr-api-changes.csv"
+api_change_document <- read.csv(url(url06), stringsAsFactors = FALSE)
 
 url_covid <- paste0(
   "https://raw.githubusercontent.com/CSSEGISandData/",
@@ -36,4 +38,5 @@ write.csv(DP03, "inputs/data/DP_03.csv", row.names = FALSE)
 write.csv(DP05, "inputs/data/DP_05.csv", row.names = FALSE)
 write.csv(covid_data, "inputs/data/covid_data.csv", row.names = FALSE)
 write.csv(election_data, "inputs/data/election_data.csv", row.names = FALSE)
-         
+write.csv(api_change_document, "inputs/data/api_change_document.csv", 
+          row.names = FALSE)

### 2023-10-29 scripts/02-data_cleaning.R
Update the cleaned data sets
Add the party each county voted for

@@ -102,7 +102,12 @@ covid_data_clean <- na.omit(covid_data_clean)
 election_data_clean <- election_data %>% 
   mutate(county = sapply(county_name, extract_county),
          state = unlist(get_state_abbreviation(state_name))) %>% 
-  select(state, county,votes_gop, votes_dem, total_votes)
+  select(state, county,votes_gop, votes_dem, total_votes) %>% 
+  rename(`votes_rep` = votes_gop)
+election_data_clean <- election_data_clean %>% 
+  mutate(party = case_when(votes_rep > votes_dem ~ "Republican",
+                           votes_rep < votes_dem ~ "Democratic",
+                           TRUE ~ "Equal"))
 election_data_clean <- na.omit(election_data_clean)
 
 # Merge the five data sets. The merged data set will be the data in this paper

### 2023-10-29 scripts/02-data_cleaning.R
Update 02-data_cleaning.R
Add infection and death rate

@@ -117,6 +117,9 @@ merged_data <- DP02_clean %>%
   inner_join(covid_data_clean, by = c("state", "county")) %>%
   inner_join(election_data_clean, by = c("state", "county"))
 
+merged_data$infection_rate <- (merged_data$cases/merged_data$total_population)*1000
+merged_data$death_rate <- (merged_data$deaths/merged_data$total_population)*1000
+
 # Export all the cleaned data sets and merged data sets
 write.csv(DP02_clean, "outputs/data/DP02_clean.csv", row.names = FALSE)
 write.csv(DP03_clean, "outputs/data/DP03_clean.csv", row.names = FALSE)

### 2023-10-29 scripts/02-data_cleaning.R
Update the voting pattern
Use 1 and 0 to indicate whether Republican won at each county

@@ -105,9 +105,9 @@ election_data_clean <- election_data %>%
   select(state, county,votes_gop, votes_dem, total_votes) %>% 
   rename(`votes_rep` = votes_gop)
 election_data_clean <- election_data_clean %>% 
-  mutate(party = case_when(votes_rep > votes_dem ~ "Republican",
-                           votes_rep < votes_dem ~ "Democratic",
-                           TRUE ~ "Equal"))
+  mutate(rep_win = case_when(votes_rep > votes_dem ~ 1,
+                             votes_rep < votes_dem ~ 0,
+                             TRUE ~ 1))
 election_data_clean <- na.omit(election_data_clean)
 
 # Merge the five data sets. The merged data set will be the data in this paper

### 2023-10-31 scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -94,7 +94,8 @@ covid_data_clean <- covid_data %>%
          `state` = Province_State,
          `cases` = Confirmed) %>%
   mutate(state = get_state_abbreviation(state)) %>% 
-  clean_names()
+  clean_names() %>% 
+  mutate(county = if_else(state == "LA", paste0(county, " Parish"), county))
 
 covid_data_clean <- na.omit(covid_data_clean)
 
@@ -105,20 +106,37 @@ election_data_clean <- election_data %>%
   select(state, county,votes_gop, votes_dem, total_votes) %>% 
   rename(`votes_rep` = votes_gop)
 election_data_clean <- election_data_clean %>% 
-  mutate(rep_win = case_when(votes_rep > votes_dem ~ 1,
+  mutate(county = if_else(state == "LA", paste0(county, " Parish"), county),
+         rep_win = case_when(votes_rep > votes_dem ~ 1,
                              votes_rep < votes_dem ~ 0,
                              TRUE ~ 1))
 election_data_clean <- na.omit(election_data_clean)
 
 # Merge the five data sets. The merged data set will be the data in this paper
-merged_data <- DP02_clean %>%
+merged_data_acs <- DP02_clean %>%
   inner_join(DP03_clean, by = c("state", "county")) %>%
-  inner_join(DP05_clean, by = c("state", "county")) %>%
-  inner_join(covid_data_clean, by = c("state", "county")) %>%
+  inner_join(DP05_clean, by = c("state", "county"))
+merged_data_acs_covid <- merged_data_acs %>% 
+  inner_join(covid_data_clean, by = c("state", "county"))
+merged_data_overall <- merged_data_acs_covid %>% 
   inner_join(election_data_clean, by = c("state", "county"))
 
-merged_data$infection_rate <- (merged_data$cases/merged_data$total_population)*1000
-merged_data$death_rate <- (merged_data$deaths/merged_data$total_population)*1000
+
+unique_rows_DP05 <- DP05_clean %>%
+  anti_join(election_data_clean, by = c("state", "county"))
+
+unique_rows_election <- election_data_clean %>%
+  anti_join(covid_data_clean, by = c("state", "county"))
+
+unique_rows_covid <- covid_data_clean %>%
+  anti_join(election_data_clean, by = c("state", "county"))
+
+
+
+merged_data_acs_covid$infection_rate <- (merged_data_acs_covid$cases/merged_data_acs_covid$total_population)*1000
[…]

### 2023-10-31 outputs/paper/paper.qmd
Update paper.qmd
First Draft

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
-author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
-date: "`r Sys.time()`"
-date-format: "D MMMM YYYY"
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+title: "aaaa"
+title-block-banner: true
+author: "Yiliu Cao"
+thanks: "Code and data from this analysis are available at: https://github.com/yiliuc/Bicycle_Thefts_2023.git"
+date: "today"
+date-format: "long"
+abstract: "sssss"
 format: pdf
+toc: true
 number-sections: true
@@
 library(palmerpenguins)
+library(here)
+library(modelsummary)
+library(leaps)
+library(gridExtra)
+library(grid)
+library(usmap)
+library(knitr)
+library(kableExtra)
+
+acs <- read.csv(here("outputs/data/merged_data_acs.csv"))
+acs_covid <- read.csv(here("outputs/data/merged_data_acs_covid.csv"))
+data <- read.csv(here("outputs/data/merged_data_overall.csv"))
+map_data <- read.csv(here("outputs/data/merged_data_overall.csv"))
+data <- data %>%
+  mutate(pctile = ntile(mean_household_income, 100),
+         party = case_when(rep_win == 1 ~ "Republican",
+                             rep_win == 0 ~ "Democratic",
+                             TRUE ~ "A"),
+         rep_rate = (votes_rep/total_votes)*1000) %>% 
+  filter(state != "AK")
+data_model <- data %>% 
+  select(-county, -state, -total_population, -females, -fips, -cases, -deaths,
+         -votes_rep, -votes_dem, -total_votes, -infection_rate, -mean_household_income)
+
+set.seed(0)
+train_indices <- sample(1:nrow(data), 0.7 * nrow(data_model))
+
+# Create the training set
+training_data <- data_model[train_indices, ]
+
+# Create the testing set
+testing_data <- data_model[-train_indices, ]
 ```
@@
 
-# Data {#sec-data}
+# Data
 
@@
 
+```{r}
+#| echo: false
+#| message: false
+#| tbl-cap: The summary of important variabls from ACS
+data %>% 
+  select(no_insurance, old_65, prop_higher_education) %>% 
+  datasummary_skim(histogram = FALSE)
+```
+
 ## CSSE at JHU
@@
 
+```{r}
+#| echo: false
+#| message: false
+#| tbl-cap: The summary of infection and death rate of COVID in US
+data %>% 
+  select(infection_rate, death_rate) %>% 
+  datasummary_skim(histogram = FALSE)
+```
+Descriptions here
+
 ## Fox News
@@
 
-## Mergin the three sources
-
 To predict the COVID death rate across counties with various socio-economic features, I will merge the the five data sets from the above three data sources by the county and state.
@@
 In addition, it also shows the demographic information such as the total population and voting pattern such as the number of votes for democratic and republican party.
-To better understand the graph, I will several graphs.
+The merged data contains three main parts of the data.
+In order to show the data better, I will draw some visualizations to show the data.
 
 ```{r}
+#| echo: false
+#| message: false
+#| label: aa
+#| fig-cap: Summary of deaths rate for counties voting for Democratic and Republican
+data  %>%
+  ggplot(mapping = aes(x = party, y = death_rate)) +
+  geom_boxplot() +
+  theme_minimal() +
+  labs(x = "Voting Party",
+       y = "Death Rate per 1000")
+```
+
+Sumamrise the above plot
+
+```{r plot-chunk, fig.width=8, fig.height=6, out.width="80%"}
+#| echo: false
+#| message: false
+#| fig-cap: Number
+
+# First plot
+plot1 <- plot_usmap(data = acs_covid, values = "death_rate", regions = "counties", exclude = c("AK", "Hawaii")) + 
+  scale_fill_continuous(low = "white", high = "blue", name = "", label = scales::comma) + 
+  labs(subtitle = "Subtitle for Plot 1") +
+  theme(legend.position = "right", plot.title = element_blank())
+
+# Second plot
+plot2 <- plot_usmap(data = data, values = "party", regions = "counties", exclude = "AK") + 
+  scale_fill_manual(values = c("Republican" = "red", "Democratic" = "blue"), 
+                    name = "Party",
+                    breaks = c("Republican", "Democratic")) + 
+  labs(subtitle = "Subtitle for Plot 2") +
+  theme(legend.position = "right", plot.title = element_blank())
+
+# Arrange the plots vertically with a main title
+grid.arrange(plot1, plot2, ncol=1)
 
@@
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+# Methods
 
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-```
+As I introduced in the introduction, this paper aim to predict the death rate of COVID in each county of US and to detect whether this correlation related to their political preferences.
+Based on the merged data we have above, I will perform a series of models, from simple to complicated, to detect which predictors are significant in predicting the death rate.
 
-Talk way more about it.
+## Model Specifics
 
-# Model
+Before we fit the models, I will split the data into training and testing data.
+The training data is used to fit all the models and they are tested on the testing model, to see which one performs betters.
 
-The goal of our modelling strategy is twofold. Firstly,...
+For each OLS regression model, the response variable will always be the death rate of COVID in each county.
+However, since the deaths rate can be attributed by various perspectives such as demographic and economics, and we also want to see whether it depends on the party they voted during the last Federal Election.
+I will create two sets of models, which the first set does not consider the political preferences but the second does.
+In addition, for each sets of model, there will be three models with each one considering only one perspective of predictors, an overall model containing all predictors and one best model based on $R^2_{adj}$ and $RMSE$.
+That said, I will fit 10 models in total along with two best models.
+To decide which model is the best one, I will use test them using the testing data, to see which one has a lower testing $RMSE$.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+In addition to OLS regression models, I will also draw the regression tree and
 
-## Model set-up
+The general form of the regression models is:
+$$Y=\beta_0+\beta_1 X_1+\cdots+\beta_p X_p+\epsilon$$
+where:
+- $Y$ represents the response variables which is the death rate.
+- $\beta_0$ represents the intercept, which is the death rate holding all variables to zero
+- $\beta_j$ represents the change of death rates with one unit increase in $X_j$
+- $\epsilon$ captures measurement errors and other discrepancies
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.\
-\[ \begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align} \] We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+## Assumptions of the regression models
+After we fit the linear regression models, it is essential to check the assumptions to ensure the accuracy of our predictions.
+There are four main assumptions which are linearity, uncorrelated error, constant variance and normality.
+For the first three assumptions, we can check it by plot the residuals with fitted values.
+However, we need to plot the Normal QQ plot to check the normality.
 
-### Model justification
+## Regression Tree and Importance Matrix
+A regression tree is a type of decision tree used specifically for regression problems, where the goal is to predict a continuous outcome variable based on one or more predictor variables.
+By using the regression tress, we can explain the decisions, identify possible events that might occur, and see potential outcomes
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+The objection function of regression tree is:
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+$$\min_{j,s} \left[  \sum_{i: x_{i,j} \leq s, x_i \in R1}
+    (y_i - \hat{y}_{R1})^2 +  \sum_{i: x_{i,j} > s, x_i \in R2}
+    (y_i - \hat{y}_{R2})^2 \right]$$
+
+The objective function is to minimize the squared error.
+The $j, s$ represents the index of the variable and threshold, respectively.
+In my regression tree, the two regions, R_1 and R_2, are split by the percentage of the White population in 2020, and the threshold here is 0.094.
+This means all the rows with the percentage of White people in 2020 less or equal to 0.094 are assigned to R1; otherwise, the rest are assigned to R2.
+In addition, the $\hat{y}_{R1}$ and $\hat{y}_{R1}$ represent the mean value of the y in each region.
+According to the regression tree, we can observe that the mean change of percentage vote for Trump is 0.157 and 0.015, which are the values of $\hat{y}_{R1}$ and $\hat{y}_{R2}$.
+
+In addition to regression trees, we can also use the importance matrix which can tell us which predictors are important in predicting the death rate for each county.
+It can help us to verify our final model.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+```{r}
+#| echo: false
+#| message: false
+#| label: first-model
+#| tbl-cap: Summary of the Best Model Without Politics Preference
+
+# Creating the Model
+best1 <- lm(death_rate ~ prop_higher_education + pctile + no_insurance +
+               males + old_85 + white_pct + black_pct, 
+            data = training_data)
+
+broom::tidy(best1) %>% 
+  mutate(predictors = c("(Intercept)", as.character(attr(terms(best1), "term.labels"))),
+         estimate = round(estimate, 2),
+         std.error = round(std.error, 2),
+         statistic = round(statistic, 2),
+         p.value = round(p.value, 3)) %>% 
+  select(predictors,
+         estimate,
+         std.error,
+         statistic,
+         p.value) %>% 
+  rename(`Predictor` = predictors,
+         `Estimate` = estimate,
+         `Standard Error` = std.error,
+         `Statistics` = statistic,
+         `P-value` = p.value) %>% 
+  kable(format = "latex")
+```
+@first-model shows the summary table for the model without political references. Using the above information, we can write the equation model:
+
+\begin{align*}
+\text{Death Rate} &=  7.23 \\
+& - 0.05 \times \text{prop\_higher\_education} \\
+& - 0.02 \times \text{pctile} \\
+& + 0.07 \times \text{no\_insurance} \\
+& - 0.06 \times \text{males} \\
+& + 0.29 \times \text{old\_85} \\
+& + 0.01 \times \text{white\_pct} \\
+& + 0.01 \times \text{black\_pct}
+\end{align*}
+
+Based on the above model, holding all variables to be zero, the expected deaths rate is about 7.2 per 1000.
+Holding other variables fixed, with one percent in the proportion of people who has at least bachelor degree, the death rate is expected to increase by 0.05.
+In addition, with income percentile increases by one, the death rate is expected to decrease by 0.02.
+Unsurprisingly, there is a negative relationship between the proportion of people with no insurance and death rate, 0.7 more people might die if the ratio of people living in a county without insurance increase by 10%.
+In terms of demographic factors, the number of death is inversely correlated with the number of males in a county.
+One percent increase in the males will result 0.06 less death.
+Moreover, it seems that old people are more likely to be died by COVID.
+That said, 0.29 more death if one percent increase in the proportion of people aged above 85.
+Lastly, one percent increase in the proportion of white and black people are both expected to have 0.01 more death.
 
@@
 #| echo: false
-#| eval: true
-#| warning: false
 #| message: false
+#| label: second-model
+#| tbl-cap: Summary of the Best Model With Politics Preference
 
+# Creating the Model
+best2 <- lm(death_rate ~ rep_rate + rep_win * prop_higher_education + pctile + 
+                 private_insurance  + no_insurance +
+                 males + old_85  + rep_win*white_pct + 
+                 black_pct, data = training_data)
+
+broom::tidy(best2) %>% 
+  mutate(predictors = c("(Intercept)", as.character(attr(terms(best2), "term.labels"))),
+         estimate = round(estimate, 2),
+         std.error = round(std.error, 2),
+         statistic = round(statistic, 2),
+         p.value = round(p.value, 3)) %>% 
+  select(predictors,
+         estimate,
+         std.error,
+         statistic,
+         p.value) %>% 
+  rename(`Predictor` = predictors,
+         `Estimate` = estimate,
+         `Standard Error` = std.error,
+         `Statistics` = statistic,
+         `P-value` = p.value) %>% 
+  knitr::kable()
 ```
 
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-```
+@second-model shows the summary of the best model without and with the political preferences. Based on them, we can write the regression question
+
+\begin{align*}
+\text{Death Rate} &=  7.7507490 \\
+& + 0.0034980 \times \text{Rep\_Rate} \\
+& - 0.4972847 \times \text{Rep\_Win} \\
+& - 0.0120727 \times \text{Prop\_Higher\_Education} \\
+& - 0.0099034 \times \text{Pctile} \\
+& - 0.0284321 \times \text{Private\_Insurance} \\
+& + 0.0281570 \times \text{No\_Insurance} \\
+& - 0.0593644 \times \text{Males} \\
+& + 0.3024402 \times \text{Old\_85} \\
+& - 0.0169198 \times \text{White\_Pct} \\
+& + 0.0130027 \times \text{Black\_Pct} \\
+& - 0.0172526 \times \text{Rep\_Win:Prop\_Higher\_Education} \\
+& + 0.0125389 \times \text{Rep\_Win:White\_Pct}
+\end{align*}
+
+The difference between the second and the first model is that the
 
@@
 
+# Appendix {-}
+## Model details
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# data_model$mean_household_income <- log(data_model$mean_household_income)
+
+social1 <- lm(death_rate ~ prop_higher_education, data = training_data)
+economic1 <- lm(death_rate ~ pctile + unemployment + no_insurance +
+                             private_insurance + public_insurance, 
+                data = training_data)
+demographic1 <- lm(death_rate ~ males + old_85 + children + old_65 +
+                                white_pct + black_pct, data = training_data)
+overall1 <- lm(death_rate ~ .-rep_win - party - rep_rate, data = training_data)
+best1 <- lm(death_rate ~ prop_higher_education + pctile + no_insurance +
+               males + old_85 + white_pct + black_pct, 
+            data = training_data)
+modelsummary(list(social1, economic1, demographic1, overall1, best1), stars=TRUE)
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| warning: false
+
+social2 <- lm(death_rate ~ rep_rate + rep_win * prop_higher_education, data = training_data)
+economic2 <- lm(death_rate ~ rep_rate + pctile + unemployment + no_insurance +
+                             rep_win*private_insurance + public_insurance, 
+                data = training_data)
+demographic2 <- lm(death_rate ~ rep_rate + males + old_85 + children + old_65 +
+                                rep_win*white_pct + black_pct, data = training_data)
+overall2 <- lm(death_rate ~ rep_rate + rep_win * prop_higher_education + pctile + 
+                 unemployment + rep_win*private_insurance + public_insurance + no_insurance +
+                 males + old_85 + children + old_65 + rep_win*white_pct + 
+                 black_pct, data = training_data)
+best2 <- lm(death_rate ~ rep_rate + rep_win * prop_higher_education + pctile + 
+                 private_insurance  + no_insurance +
+                 males + old_85  + rep_win*white_pct + 
+                 black_pct, data = training_data)
+modelsummary(list(social2, economic2, demographic2, overall2, best2), stars=TRUE)
+```
+
 \newpage

### 2023-10-31 outputs/paper/paper.qmd
Finalize all the models

@@
                              TRUE ~ "A"),
-         rep_rate = (votes_rep/total_votes)*1000) %>% 
+         rep_rate = (votes_rep/total_votes)*1000,
+         high_income = case_when(pctile > 50 ~ 1,
+                                 TRUE ~ 0)) %>% 
   filter(state != "AK")
@@
   select(-county, -state, -total_population, -females, -fips, -cases, -deaths,
-         -votes_rep, -votes_dem, -total_votes, -infection_rate, -mean_household_income)
+         -votes_rep, -votes_dem, -total_votes, -infection_rate, -mean_household_income, -old_65)
 
@@
 
-```{r plot-chunk, fig.width=8, fig.height=6, out.width="80%"}
+```{r plot-chunk, fig.width=8, fig.height=6, out.width="100%"}
 #| echo: false
@@
 # Second plot
-plot2 <- plot_usmap(data = data, values = "party", regions = "counties", exclude = "AK") + 
+plot2 <- plot_usmap(data = data, values = "party", regions = "counties", exclude = "AK", "Hawaii") + 
   scale_fill_manual(values = c("Republican" = "red", "Democratic" = "blue"), 
@@
 
+@first-model shows the summary table for the model without political references. Using the above information, we can write the equation model:
+
 ```{r}
-#| echo: false
-#| message: false
-#| label: first-model
-#| tbl-cap: Summary of the Best Model Without Politics Preference
+model1 <- lm(death_rate ~ prop_higher_education + pctile + no_insurance + private_insurance + males + old_85 + white_pct + black_pct, data = training_data)
+summary(model1)
+```
 
-# Creating the Model
-best1 <- lm(death_rate ~ prop_higher_education + pctile + no_insurance +
-               males + old_85 + white_pct + black_pct, 
-            data = training_data)
-
-broom::tidy(best1) %>% 
-  mutate(predictors = c("(Intercept)", as.character(attr(terms(best1), "term.labels"))),
-         estimate = round(estimate, 2),
-         std.error = round(std.error, 2),
-         statistic = round(statistic, 2),
-         p.value = round(p.value, 3)) %>% 
-  select(predictors,
-         estimate,
-         std.error,
-         statistic,
-         p.value) %>% 
-  rename(`Predictor` = predictors,
-         `Estimate` = estimate,
-         `Standard Error` = std.error,
-         `Statistics` = statistic,
-         `P-value` = p.value) %>% 
-  kable(format = "latex")
+```{r}
+model2 <- lm(death_rate ~ rep_rate + prop_higher_education + pctile + 
+                 private_insurance  + no_insurance +
+                 males + old_85  + white_pct + 
+                 black_pct, data = training_data)
+summary(model2)
 ```
-@first-model shows the summary table for the model without political references. Using the above information, we can write the equation model:
 
@@
 
+@second-model shows the summary of the best model without and with the political preferences. Based on them, we can write the regression question
+
+\begin{align*}
+\text{Death Rate} &=  7.75 \\
+& + 0.0034980 \times \text{Rep\_Rate} \\
+& - 0.4972847 \times \text{Rep\_Win} \\
+& - 0.0120727 \times \text{Prop\_Higher\_Education} \\
+& - 0.0099034 \times \text{Pctile} \\
+& - 0.0284321 \times \text{Private\_Insurance} \\
+& + 0.0281570 \times \text{No\_Insurance} \\
+& - 0.0593644 \times \text{Males} \\
+& + 0.3024402 \times \text{Old\_85} \\
+& - 0.0169198 \times \text{White\_Pct} \\
+& + 0.0130027 \times \text{Black\_Pct} \\
+& - 0.0172526 \times \text{Rep\_Win:Prop\_Higher\_Education} \\
+& + 0.0125389 \times \text{Rep\_Win:White\_Pct}
+\end{align*}
+
+The difference between the second and the first model is that the second model includes the political preference for each county as one predictor to predict the death rate due to COVID.
+Holding other variables, we can see that every 100 more votes for republican party, 0.3 more people are expected to dead.
+In addition, if the county votes for the Republican, then it is expected to have
+
+# Discussion
+
+## The impact of political preference on death rate is more significant than expected.
+
+If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+
+## However, the socio-economic variables especially the private insurance, also play an important role.
+
+## The importantance matrix matches our models
+
+## Weaknesses and next steps
+
+Weaknesses and next steps should also be included.
+
+\newpage
+
+# Appendix {-}
+
+## Model details
+
+aaa
+
+ojnoj
+
+klmko
+
+```{r}
+#| echo: false
+#| message: false
+#| label: first-model
+#| tbl-cap: Summary of the Best Model Without Politics Preference
+
+# Creating the Model
+best1 <- lm(death_rate ~ prop_higher_education + pctile + no_insurance + private_insurance + males + old_85 + white_pct + black_pct, data = training_data)
+
+broom::tidy(best1) %>% 
+  mutate(predictors = c("(Intercept)", as.character(attr(terms(best1), "term.labels"))),
+         estimate = round(estimate, 2),
+         std.error = round(std.error, 2),
+         statistic = round(statistic, 2),
+         p.value = round(p.value, 3)) %>% 
+  select(predictors,
+         estimate,
+         std.error,
+         statistic,
+         p.value) %>% 
+  rename(`Predictor` = predictors,
+         `Estimate` = estimate,
+         `Standard Error` = std.error,
+         `Statistics` = statistic,
+         `P-value` = p.value) %>% 
+  kable(format = "latex")
+```
+
+j jnjn
+
 ```{r}
@@
 # Creating the Model
-best2 <- lm(death_rate ~ rep_rate + rep_win * prop_higher_education + pctile + 
+best2 <- lm(death_rate ~ rep_rate + prop_higher_education + pctile + 
                  private_insurance  + no_insurance +
-                 males + old_85  + rep_win*white_pct + 
+                 males + old_85  + white_pct + 
                  black_pct, data = training_data)
@@
 
-@second-model shows the summary of the best model without and with the political preferences. Based on them, we can write the regression question
-
-\begin{align*}
-\text{Death Rate} &=  7.7507490 \\
-& + 0.0034980 \times \text{Rep\_Rate} \\
-& - 0.4972847 \times \text{Rep\_Win} \\
-& - 0.0120727 \times \text{Prop\_Higher\_Education} \\
-& - 0.0099034 \times \text{Pctile} \\
-& - 0.0284321 \times \text{Private\_Insurance} \\
-& + 0.0281570 \times \text{No\_Insurance} \\
-& - 0.0593644 \times \text{Males} \\
-& + 0.3024402 \times \text{Old\_85} \\
-& - 0.0169198 \times \text{White\_Pct} \\
-& + 0.0130027 \times \text{Black\_Pct} \\
-& - 0.0172526 \times \text{Rep\_Win:Prop\_Higher\_Education} \\
-& + 0.0125389 \times \text{Rep\_Win:White\_Pct}
-\end{align*}
-
-The difference between the second and the first model is that the
-
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-# Appendix {-}
-## Model details
 ```{r}
 #| echo: false
-#| eval: true
-#| warning: false
 #| message: false
+#| tbl-cap: Summary of the set of models with political preferences
 
@@
                 data = training_data)
-demographic1 <- lm(death_rate ~ males + old_85 + children + old_65 +
+demographic1 <- lm(death_rate ~ males + old_85 + children +
                                 white_pct + black_pct, data = training_data)
 overall1 <- lm(death_rate ~ .-rep_win - party - rep_rate, data = training_data)
-best1 <- lm(death_rate ~ prop_higher_education + pctile + no_insurance +
-               males + old_85 + white_pct + black_pct, 
-            data = training_data)
+best1 <- lm(death_rate ~ prop_higher_education + pctile + no_insurance + private_insurance + males + old_85 + white_pct + black_pct, data = training_data)
 modelsummary(list(social1, economic1, demographic1, overall1, best1), stars=TRUE)
@@
 #| echo: false
-#| eval: true
+#| message: false
 #| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
+#| tbl-cap: Summary of the set of models with political preferences
 
-social2 <- lm(death_rate ~ rep_rate + rep_win * prop_higher_education, data = training_data)
+social2 <- lm(death_rate ~ rep_rate + prop_higher_education, data = training_data)
 economic2 <- lm(death_rate ~ rep_rate + pctile + unemployment + no_insurance +
-                             rep_win*private_insurance + public_insurance, 
+                             private_insurance + public_insurance, 
                 data = training_data)
-demographic2 <- lm(death_rate ~ rep_rate + males + old_85 + children + old_65 +
-                                rep_win*white_pct + black_pct, data = training_data)
-overall2 <- lm(death_rate ~ rep_rate + rep_win * prop_higher_education + pctile + 
-                 unemployment + rep_win*private_insurance + public_insurance + no_insurance +
-                 males + old_85 + children + old_65 + rep_win*white_pct + 
+demographic2 <- lm(death_rate ~ rep_rate + males + old_85 + children +
+                                white_pct + black_pct, data = training_data)
+overall2 <- lm(death_rate ~ rep_rate + prop_higher_education + pctile + 
+                 unemployment + private_insurance + public_insurance + no_insurance +
+                 males + old_85 + children + white_pct + 
                  black_pct, data = training_data)
-best2 <- lm(death_rate ~ rep_rate + rep_win * prop_higher_education + pctile + 
+best2 <- lm(death_rate ~ rep_rate + prop_higher_education + pctile + 
                  private_insurance  + no_insurance +
-                 males + old_85  + rep_win*white_pct + 
+                 males + old_85  + white_pct + 
                  black_pct, data = training_data)

### 2023-11-01 outputs/paper/paper.qmd
Update the first draft

--- 2023-10-24 outputs/paper/paper.qmd
@@
 library(tidyverse)
-library(palmerpenguins)
+library(here)
+library(modelsummary)
+library(gridExtra)
+library(usmap)
+library(knitr)
+library(kableExtra)
+library(purrr)
+
+acs <- read.csv(here("outputs/data/merged_data_acs.csv"))
+acs_covid <- read.csv(here("outputs/data/merged_data_acs_covid.csv"))
+data <- read.csv(here("outputs/data/merged_data_overall.csv"))
+map_data <- read.csv(here("outputs/data/merged_data_overall.csv"))
+election_data <- read.csv(here("outputs/data/election_data_clean.csv"))
+data <- data %>%
+  mutate(pctile = ntile(mean_household_income, 100),
+         party = case_when(rep_win == 1 ~ "Republican",
+                             rep_win == 0 ~ "Democratic",
+                             TRUE ~ "A"),
+         high_income = case_when(pctile > 50 ~ 1,
+                                 TRUE ~ 0),
+         age_85 = old_85,
+         high_education = prop_higher_education,
+         white = white_pct,
+         black = black_pct) %>% 
+  filter(state != "AK")
+data_model <- data %>% 
+  select(-county, -state, -total_population, -females, -fips, -cases, -deaths,
+         -votes_rep, -votes_dem, -total_votes, -infection_rate, -mean_household_income, -old_65)
+
+set.seed(0)
+train_indices <- sample(1:nrow(data), 0.7 * nrow(data_model))
+
+# Create the training set
+training_data <- data_model[train_indices, ]
+
+# Create the testing set
+testing_data <- data_model[-train_indices, ]
 ```
@@
 
-## American Community Survey (ACS)
+## Source of Data
+
+This paper will use five data sets from three different sources. They are shown below.
+
+### American Community Survey (ACS)
 
@@
 
-## CSSE at JHU
+```{r}
+#| echo: false
+#| message: false
+#| tbl-cap: The summary of important variabls from ACS
+data %>% 
+  select(no_insurance, age_85, high_education) %>% 
+  datasummary_skim(histogram = FALSE)
+```
+
+### CSSE at JHU
+
 The CSSE at JHU collects the worldwide COVID data since 2020 to 2023.
@@
 
-## Fox News
+```{r}
+#| echo: false
+#| message: false
+#| tbl-cap: The summary of infection and death rate of COVID in US
+data %>% 
+  select(infection_rate, death_rate) %>% 
+  datasummary_skim(histogram = FALSE)
+```
+
+Descriptions here
+
+### Fox News
 
@@
 
-## Mergin the three sources
+## Data Cleaning
 
-To predict the COVID death rate across counties with various socio-economic features, I will merge the the five data sets from the above three data sources by the county and state.
-In the merged data set, each row represent a US county and provides the number of COVID cases and deaths.
-In addition, it also shows the demographic information such as the total population and voting pattern such as the number of votes for democratic and republican party.
-To better understand the graph, I will several graphs.
+All data was collected and cleaned using the statistical programming language `R` \[citation\].
+The packages used are `tidyverse` \[citation\], `dplyr` \[citation\], `stringr` \[citation\] and `janitor` \[citation\].
+Besides, `here` \[citation\], `modelsummary` \[citation\], `girdExtra` \[citation\], `usmap` \[citation\], `knitr` \[citation\], `kableExtra` \[citation\] are used to analyse the data.
+
+The procedure to clean the three sources of data are different.
+For all the ACS data downloaded using ACS API \[citation\], referring to the codebook published by ACS \[citation\], I only extract the necessary variables.
+Besides, I clean the name for each county and convert the state name to its abbreviation form.
+In terms of the COVID data from JHU, I only select the cases and deaths and filter only the counties in US.
+Lastly for the election data, I extract the county name and also transform the state name to its abbreviation form.
+I also calculate the number of people voting for Republican party per 1000 voters.
+
+After cleaned each data set, I merge the the five data sets from the above three data sources by the county and state.
+Based on the merged data set, I calculate the infection and death number per 1000 residents in each county.
+Therefore, each row in the merged data represent a US county and provides the the demographic information such as the total population.
+In addition, the data also contains the voting pattern in each county such as the number of votes for democratic and republican party.
 
 ```{r}
-
+#| echo: false
+#| message: false
+#| label: tbl-summary_variables
+#| tbl-cap: Summary of important variables in the model
+data.frame(variable = c("High education attainment",
+                        "Income Percentile",
+                        "People with no insurance",
+                        "People with private insurance",
+                        "Males",
+                        "People aged above 85",
+                        "White people",
+                        "Black people",
+                        "People supporting Republican"),
+           coded_name = c("high_education",
+                          "IncomePctile",
+                          "no_insurance",
+                          "private_insurance",
+                          "males",
+                          "age_85",
+                          "white",
+                          "black",
+                          "rep_rate"),
+           description = c("The proportion of local residents with at least Bachelor degree",
+                           "The mean household income percentile representing the income level for each county",
+                           "The proportion of local residents without insurance",
+                           "The proportion of local residents with private insurance",
+                           "The proportion of male residents",
+                           "The proportion of people aged above 85",
+                           "The proportion of white people",
+                           "The proportion of black people",
+                           "Number of voters supporting Republican by 1000")) %>% 
+  rename(`Variable` = variable,
+         `Encoded Name` = coded_name,
+         `Description` = description) %>% 
+  kable(booktabs = TRUE) %>% 
+  column_spec(column = 3, width = "6cm")
 ```
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+The @tbl-summary_variables summarizes all the important variables in this paper.
+It contains the social characteristics such as high education attainment and economic factors such as income percentile.
+Moreover, the demographic information and political preference for each county are also included
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| echo: false
+#| message: false
+#| label: fig-boxplot
+#| fig-cap: Summary of deaths rate for counties voting for Democratic and Republican
+data  %>%
+  ggplot(mapping = aes(x = party, y = death_rate)) +
+  geom_boxplot() +
+  theme_minimal() +
+  labs(x = "Voting Party",
+       y = "Death Rate per 1000")
+```
+
+To understand the data better, @fig-boxplot shows the the distribution of death rate for the counties voting for the two parties.
+We can observe that difference in political preference results differences in death rate.
+Counties voting for republican has a relative higher death rate than the counties voting for the Democratic party.
+In addition, there have outliers for both groups.
+
+```{r plot-chunk, fig.width=8, fig.height=6, out.width="100%"}
 #| echo: false
@@
 #| message: false
+#| label: fig-map
+#| fig-cap: Election results of 2020 Federal Election and the corresponding death rate for each US county
+state_boundaries <- usmap::us_map(regions = "states", 
+                                  exclude = c("AK", "Hawaii"))
+plot1 <- plot_usmap(data = data, values = "rep_rate", regions = "counties", exclude = c("AK", "Hawaii")) + 
+  scale_fill_gradient2(name = "Voting for the Republican %", 
+                       low = "blue", mid = "white", high = "red",
+                       midpoint = 500, 
+                       breaks = c(0, 500, 1000)) + 
+  geom_polygon(data = state_boundaries, 
+               aes(x = x, y = y, group = group), 
+               fill = NA, 
+               color = "black", size = 0.8) +  
+  labs(subtitle = "The voting patter for each US county") +
+  theme(plot.subtitle = element_text(size = 12,
+                                     hjust = 0.5),
+        legend.position = "right", plot.title = element_blank())
+
+plot2 <- plot_usmap(data = data, values = "death_rate", regions = "counties", exclude = c("AK", "Hawaii")) + 
+  scale_fill_gradient2(low = "white", high = "blue", 
+                        limits = c(0, 8),
+                        breaks = c(0, 8),
+                        name = "Death per 1000",
+                       oob = scales::squish) + 
+  geom_polygon(data = state_boundaries, 
+               aes(x = x, y = y, group = group), 
+               fill = NA, 
+               color = "black", size = 0.8) +  
+  labs(subtitle = "Death of COVID per 1000 ") +
+  theme(plot.subtitle = element_text(size = 12,
+                                     hjust = 0.5),
+        legend.position = "right", plot.title = element_blank())
+
+# Arrange the plots vertically with a main title
+library(cowplot)
+
+plot_grid(plot1, plot2, ncol = 1, align = "v", rel_heights = c(1,1))
 ```
 
-Talk way more about it.
+To understand how the death rate correlates with the political preference, @fig-map compares the support for Republican with corresponding death rate.
+From the graph, we can observe that the dark color corresponds to the dark blue, meaning that the counties voting for Republican seem to have a higher death rate than the counties voting for Democratic party.
+This pattern is especially significant for the central states.
 
-# Model
+## Data limitations
 
-The goal of our modelling strategy is twofold. Firstly,...
+Our data combines five data sets from three sources. However, it still have limitations which may affect the analyses, reducing the accuracy of our predictions.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+First of all, there is lack of COVID information in certain states, especially for the state "Utah." We can observe this from @fig-map which some counties shows a color of grey.
+Due to this lackness, the merged data set will also lack the information on these counties even though other data sets have that information.
+The accuracy of the models may be also affected.
+In addition, the data we have may lack of spread, meaning that there could have other information which may also affect the death rate but are not included in this paper such as the hospital capacity.
+Our analyse may be less convincing.
 
-## Model set-up
+# Methods
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.\
-\[ \begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align} \] We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+As I introduced in the introduction, this paper aim to predict the death rate of COVID in each county of US and to detect whether this correlation related to their political preferences.
+Based on the merged data we have above, I will perform a series of models, from simple to complicated, to detect which predictors are significant in predicting the death rate.
 
-### Model justification
+## Model Specifics
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+Before we fit the models, I will split the data into training and testing data.
+The training data is used to fit all the models and they are tested on the testing model, to see which one performs betters.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+For each OLS regression model, the response variable will always be the death rate of COVID in each county.
+However, since the deaths rate can be attributed by various perspectives such as demographic and economics, and we also want to see whether it depends on the party they voted during the last Federal Election.
+I will create two sets of models, which the first set does not consider the political preferences but the second does.
+In addition, for each sets of model, there will be three models with each one considering only one perspective of predictors, an overall model containing all predictors and one best model based on $R^2_{adj}$ and $RMSE$.
+That said, I will fit 10 models in total along with two best models.
+To decide which model is the best one, I will use test them using the testing data, to see which one has a lower testing $RMSE$.
+
+In addition to OLS regression models, I will also draw the regression tree and
+
+The general form of the regression models is: $$Y=\beta_0+\beta_1 X_1+\cdots+\beta_p X_p+\epsilon$$ where: - $Y$ represents the response variables which is the death rate.
+- $\beta_0$ represents the intercept, which is the death rate holding all variables to zero - $\beta_j$ represents the change of death rates with one unit increase in $X_j$ - $\epsilon$ captures measurement errors and other discrepancies
+
+## Assumptions of the regression models
+
+After we fit the linear regression models, it is essential to check the assumptions to ensure the accuracy of our predictions.
+There are four main assumptions which are linearity, uncorrelated error, constant variance and normality.
+For the first three assumptions, we can check it by plot the residuals with fitted values.
+However, we need to plot the Normal QQ plot to check the normality.
+
+## Regression Tree and Importance Matrix
+
+A regression tree is a type of decision tree used specifically for regression problems, where the goal is to predict a continuous outcome variable based on one or more predictor variables.
+By using the regression tress, we can explain the decisions, identify possible events that might occur, and see potential outcomes
+
+The objection function of regression tree is:
+
+$$\min_{j,s} \left[  \sum_{i: x_{i,j} \leq s, x_i \in R1}
+    (y_i - \hat{y}_{R1})^2 +  \sum_{i: x_{i,j} > s, x_i \in R2}
+    (y_i - \hat{y}_{R2})^2 \right]$$
+
+The objective function is to minimize the squared error.
+The $j, s$ represents the index of the variable and threshold, respectively.
+In my regression tree, the two regions, R_1 and R_2, are split by the percentage of the White population in 2020, and the threshold here is 0.094.
+This means all the rows with the percentage of White people in 2020 less or equal to 0.094 are assigned to R1; otherwise, the rest are assigned to R2.
+In addition, the $\hat{y}_{R1}$ and $\hat{y}_{R1}$ represent the mean value of the y in each region.
+According to the regression tree, we can observe that the mean change of percentage vote for Trump is 0.157 and 0.015, which are the values of $\hat{y}_{R1}$ and $\hat{y}_{R2}$.
+
+In addition to regression trees, we can also use the importance matrix which can tell us which predictors are important in predicting the death rate for each county.
+It can help us to verify our final model.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+## Death rate without considering the political preferences
 
@@
 #| echo: false
-#| eval: true
-#| warning: false
 #| message: false
+#| label: tbl-first_model
+#| tbl-cap: Summary of the Best Model Without Politics Preference
 
+# Creating the Model
+best1 <- lm(death_rate ~ high_education + pctile + no_insurance + private_insurance + males + age_85 + white + black, data = training_data)
+
+broom::tidy(best1) %>% 
+  mutate(predictors = c("(Intercept)", as.character(attr(terms(best1), "term.labels"))),
+         estimate = round(estimate, 3),
+         std.error = round(std.error, 3),
+         statistic = round(statistic, 3),
+         p.value = round(p.value, 3)) %>% 
+  select(predictors,
+         estimate,
+         std.error,
+         statistic,
+         p.value) %>% 
+  rename(`Predictor` = predictors,
+         `Estimate` = estimate,
+         `Standard Error` = std.error,
+         `Statistics` = statistic,
+         `P-value` = p.value) %>% 
+  kable(booktabs = TRUE)
+```
+
+@tbl-first_model shows the summary table for the model without political references. Using the above information, we can write the equation model:
+
+```{=tex}
+\begin{align*}
+\text{Death Rate} &=  7.521 \\
+& - 0.051 \times \text{prop\_higher\_education} \\
+& - 0.012 \times \text{IncomePctile} \\
+& + 0.066 \times \text{no\_insurance} \\
+& - 0.014 \times \text{private\_insurance} \\
+& - 0.055 \times \text{males} \\
+& + 0.303 \times \text{old\_85} \\
+& + 0.011 \times \text{white\_pct} \\
+& + 0.013 \times \text{black\_pct}
+\end{align*}
+```
+Based on the above model, holding all variables to be zero, the expected deaths rate is about 7.2 per 1000.
+Holding other variables fixed, with one percent in the proportion of people who has at least bachelor degree, the death rate is expected to increase by 0.05.
+In addition, with income percentile increases by one, the death rate is expected to decrease by 0.02.
+Unsurprisingly, there is a negative relationship between the proportion of people with no insurance and death rate, 0.7 more people might die if the ratio of people living in a county without insurance increase by 10%.
+In terms of demographic factors, the number of death is inversely correlated with the number of males in a county.
+One percent increase in the males will result 0.06 less death.
+Moreover, it seems that old people are more likely to be died by COVID.
+That said, 0.29 more death if one percent increase in the proportion of people aged above 85.
+Lastly, one percent increase in the proportion of white and black people are both expected to have 0.01 more death.
+
+## Death rate with considering the plotical preference
+
+```{r}
+#| echo: false
+#| message: false
+#| label: tbl-second_model
+#| tbl-cap: Summary of the Best Model With Politics Preference
+
+# Creating the Model
+best2 <- lm(death_rate ~ rep_rate + high_education + pctile + 
+                 private_insurance  + no_insurance +
+                 males + age_85  + white + 
+                 black, data = training_data)
+
+broom::tidy(best2) %>% 
+  mutate(predictors = c("(Intercept)", as.character(attr(terms(best2), "term.labels"))),
+         estimate = round(estimate, 3),
+         std.error = round(std.error, 3),
+         statistic = round(statistic, 3),
+         p.value = round(p.value, 3)) %>% 
+  select(predictors,
+         estimate,
+         std.error,
+         statistic,
+         p.value) %>% 
+  rename(`Predictor` = predictors,
+         `Estimate` = estimate,
+         `Standard Error` = std.error,
+         `Statistics` = statistic,
+         `P-value` = p.value) %>% 
+  kable(booktabs = TRUE)
+```
+
+@tbl-second_model shows the summary of the best model without and with the political preferences. Based on them, we can write the regression question.
+
+```{=tex}
+\begin{align*}
+\text{Death Rate} &=  7.596 \\
+& + 0.003 \times \text{Rep\_Rate} \\
+& - 0.024 \times \text{High\_Education} \\
+& - 0.010 \times \text{Pctile} \\
+& + 0.027 \times \text{No\_Insurance} \\
+& - 0.028 \times \text{Private\_Insurance} \\
+& - 0.059 \times \text{Males} \\
+& + 0.308 \times \text{Age\_85} \\
+& - 0.008 \times \text{White} \\
+& + 0.013 \times \text{Black}
+\end{align*}
+```
+The difference between the second and the first model is that the second model includes the political preference for each county as one predictor to predict the death rate due to COVID.
+Holding other variables, we can see that every 100 more votes per 1000 for republican party, 0.3 more people are expected to dead.
+
+## Comparing the testing error
+
+The above two models are fitted using the training data.
+Comparing the training RMSE from @tbl-all_models_1 and @tbl-all_models_2, we can find that the model containing the political preference has a lower training RMSE.
+However, in order to test the prediction performance of the models we have varying data, I will use testing data to find the test error for the two models above.
+
+```{r}
+#| echo: false
+#| message: false
+#| label: tbl-test_error
+#| tbl-cap: The testing error of the above two models
+
+calculate_testing_error <- function(model, testing_data, outcome) {
+  # Predict the outcomes on the testing data
+  predictions <- predict(model, newdata = testing_data)
+
+  # Compute the residuals
+  residuals <- predictions - testing_data[[outcome]]
+
+  # Compute the MSE
+  mse <- mean(residuals^2)
+
+  return(mse)
+}
+
+data.frame(model = c("Model 1", "Model 2"),
+           error = c(calculate_testing_error(best1, testing_data, "death_rate"),
+                     calculate_testing_error(best2, testing_data, "death_rate"))) %>% 
+  rename(`Model` = model,
+         `Testinf Error` = error) %>% 
+  kable(booktabs = TRUE)
+```
+
+@tbl-test_error shows the testing error of the above two models.
+We can observe that the errors for the two models are 1.769733 and 1.675969.
+Therefore, the model containing the political preferences has a lower expecter MSE, hence it has better predictions than another one.
+
+# Discussion
+
+To sum up, the second model which contains the political reference is more appropriate than another one, representing that the party each county supported during the last election may change the attitude that people face with COVID and hence results differences in death rate.
+In addition to that, it is also important to consider the rest variables while predicting the death rate.
+
+## The impact of political preference on death rate is more significant than expected, it may overcome the change of income levels.
+
+From the @tbl-second_model, we can observe that both income percentile and the support of Republican both are significant in predicting the death rate of COVID, but with an inverse direction of their coefficients.
+More people voting for Republican party is expected to increase the number of deaths due to COVID in this county and a wealthier counties usually have a less COVID deaths.
+However, if we compare their values of coefficients, we can find that the increase in death by every three more people in 1000 voting for Republican party is equivalent to decrease in death rate by one increase in the income percentile.
+In other words, counties with the lowest income levels might be the worst areas sacrificing from COVID if they support the Republican party.
+
+This is a surprising but also horrible observation as Donald Trump is especially in favor of poor people.
+Eight out of the top 10 counties with highest poverty rate voted for Donald Trump.
+Referring to @fig-map, we can clearly see that the above 80% states in the poorer states, middle US, had voted for Trump.
+Conversely, above 80% wealthier coastal states, such as California and New York, had voted for The Democratic party.
+These both verify our observations from the model.
+What Trump's votes does not know is that their support may bring more deaths to their counties.
+
+But why, why the counties with higher income levels are less likely to vote for the Republican party and hence results in a lower death rate?
+One thing that can explain this pattern from our model is the proportion of people that have higher education level.
+According to @tbl-second_model, we can find that there is an inverse relationship between the death rate and the number of residents with high education level.
+Its value of coefficient is much higher than the previous two, indicating that education level is an essential factor in predicting the death rate.
+
+```{r}
+#| echo: false
+#| message: false
+#| label: fig-education
+#| fig-cap: The testing error of the above two models
+
+breakpoints <- quantile(data$high_education, probs = c(0, 1/3, 2/3, 1))
+data$education <- cut(data$high_education, breaks = breakpoints, 
+                      labels = c("Low", "Medium", "High"), include.lowest = TRUE)
+
+ggplot(data, aes(x = high_education, y = death_rate)) +
+  geom_point() +
+  labs(x = "Income Percentile", y = "Death Rate", title = "Death Rate by Education Level and Income Percentile") +
+  theme_minimal()
 ```
@@
 ```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
+summarized_data <- data %>%
+  mutate(bin = cut(pctile, breaks = seq(0, 100, by = 5), include.lowest = TRUE, labels = seq(2.5, 97.5, by = 5))) %>%
+  group_by(bin) %>%
+  summarise(
+    avg_death_rate = mean(death_rate, na.rm = TRUE),
+    avg_education_level = mean(high_education, na.rm = TRUE)
+  )
+
+# Create the plot with dual y-axis
+ggplot(summarized_data, aes(x = as.numeric(as.character(bin)))) +
+  geom_line(aes(y = avg_death_rate, color = "Average Death Rate"), size = 1) +
+  geom_line(aes(y = avg_education_level/2, color = "Average Education Level"), size = 1) + # Adjust the division for better visualization
+  scale_y_continuous(sec.axis = sec_axis(~.*2, name = "Average Education Level")) + # Adjust the multiplier to reverse the transformation
+  labs(x = "Income Percentile", y = "Average Death Rate", color = "Legend") +
+  theme_minimal()
 ```
 
-# Discussion
+Education reflects this
 
-## First discussion point {#sec-first-point}
+## Health insurance iss also an important factor in predicting the death rate of COVID.
+In addition to the political preferences, the healthcare insurance is also significant in predicting the death rate for each county.
+Based on @tbl-second_model, it seems that both `no_insurance` and `private_insurance` are also significant in predicting the death rate for each county but they shows an inverse relationship on the death rate.
+In addition, the values of these two coefficients are almost same, meaning that the effects on death rate between them can be offset.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+However, t
 
-## Second discussion point
-
-## Third discussion point
+## The importantance matrix matches our models
 
@@
 
+# Appendix {.unnumbered}
+
+## Model details
+
+aaa
+
+ojnoj
+
+klmko
+
+j jnjn
+
+```{r}
+#| echo: false
+#| message: false
+#| label: tbl-all_models_1
+#| tbl-cap: Summary of the set of models with political preferences
+
+# data_model$mean_household_income <- log(data_model$mean_household_income)
+
+social1 <- lm(death_rate ~ high_education, data = training_data)
+economic1 <- lm(death_rate ~ pctile + unemployment + no_insurance +
+                             private_insurance + public_insurance, 
+                data = training_data)
+demographic1 <- lm(death_rate ~ males + age_85 + children +
+                                white + black, data = training_data)
+overall1 <- lm(death_rate ~ .-rep_win - party - rep_rate - high_income, data = training_data)
+best1 <- lm(death_rate ~ high_education + pctile + no_insurance + private_insurance + males + age_85 + white + black, data = training_data)
+modelsummary(list(social1, economic1, demographic1, overall1, best1), stars=TRUE)
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| label: tbl-all_models2
+#| tbl-cap: Summary of the set of models with political preferences
+
+social2 <- lm(death_rate ~ rep_rate + high_education, data = training_data)
+economic2 <- lm(death_rate ~ rep_rate + pctile + unemployment + no_insurance +
+                             private_insurance + public_insurance, 
+                data = training_data)
+demographic2 <- lm(death_rate ~ rep_rate + males + age_85 + children +
+                                white + black, data = training_data)
+overall2 <- lm(death_rate ~ rep_rate + high_education + pctile + 
+                 unemployment + private_insurance + public_insurance + no_insurance +
+                 males + age_85 + children + white + 
+                 black, data = training_data)
+best2 <- lm(death_rate ~ rep_rate + high_education + pctile + 
+                 private_insurance  + no_insurance +
+                 males + age_85  + white + 
+                 black, data = training_data)
+modelsummary(list(social2, economic2, demographic2, overall2, best2), stars=TRUE)
+```
+
 \newpage
@@
 https://news.un.org/en/story/2023/05/1136367 https://www.aljazeera.com/news/2023/5/11/three-years-1-1-million-deaths-covid-emergency-ending-in-us
+https://wisevoter.com/state-rankings/poorest-states/

### 2023-11-01 scripts/02-data_cleaning.R
Update the cleaned data

@@ -109,7 +109,8 @@ election_data_clean <- election_data_clean %>%
   mutate(county = if_else(state == "LA", paste0(county, " Parish"), county),
          rep_win = case_when(votes_rep > votes_dem ~ 1,
                              votes_rep < votes_dem ~ 0,
-                             TRUE ~ 1))
+                             TRUE ~ 1),
+         rep_rate = (votes_rep/total_votes)*1000,)
 election_data_clean <- na.omit(election_data_clean)
 
 # Merge the five data sets. The merged data set will be the data in this paper
@@ -122,14 +123,17 @@ merged_data_overall <- merged_data_acs_covid %>%
   inner_join(election_data_clean, by = c("state", "county"))
 
 
-unique_rows_DP05 <- DP05_clean %>%
-  anti_join(election_data_clean, by = c("state", "county"))
-
-unique_rows_election <- election_data_clean %>%
-  anti_join(covid_data_clean, by = c("state", "county"))
-
-unique_rows_covid <- covid_data_clean %>%
-  anti_join(election_data_clean, by = c("state", "county"))
+# unique_rows_DP05 <- DP05_clean %>%
+#   anti_join(covid_data_clean, by = c("state", "county"))
+# 
+# unique_rows_election <- election_data_clean %>%
+#   anti_join(covid_data_clean, by = c("state", "county"))
+# 
+# unique_rows_election <- election_data_clean %>%
+#   anti_join(covid_data_clean, by = c("state", "county"))
+# 
+# unique_rows_covid <- covid_data_clean %>%
+#   anti_join(election_data_clean, by = c("state", "county"))
 
 
 

### 2023-11-01 outputs/paper/paper.qmd
Finish the first draft!

@@
 
-## The impact of political preference on death rate is more significant than expected, it may overcome the change of income levels.
+## Politic preference to income levels: wealthier people are less likely to vote for the Republican, perhaps they are more educated.
 
@@
 #| message: false
-#| label: fig-education
-#| fig-cap: The testing error of the above two models
-
-breakpoints <- quantile(data$high_education, probs = c(0, 1/3, 2/3, 1))
-data$education <- cut(data$high_education, breaks = breakpoints, 
-                      labels = c("Low", "Medium", "High"), include.lowest = TRUE)
-
-ggplot(data, aes(x = high_education, y = death_rate)) +
-  geom_point() +
-  labs(x = "Income Percentile", y = "Death Rate", title = "Death Rate by Education Level and Income Percentile") +
-  theme_minimal()
-```
-
-```{r}
+#| label: fig-income_education
+#| fig-cap: The change of average death rate across different education levels
 summarized_data <- data %>%
@@
     avg_death_rate = mean(death_rate, na.rm = TRUE),
-    avg_education_level = mean(high_education, na.rm = TRUE)
+    avg_education_prop = mean(high_education, na.rm = TRUE)
+  )
+
+# Create the plot with dual y-axis
+# Create the plot with dual y-axis
+ggplot(summarized_data, aes(x = as.numeric(as.character(bin)))) +
+  geom_line(aes(y = avg_death_rate, color = "Avg. Death Rate"), size = 1) +
+  geom_line(aes(y = avg_education_prop/5, color = "Avg. High-Education Rates"), size = 1) + 
+  scale_y_continuous(sec.axis = sec_axis(~.*5, name = "Avg. High-Education Rates")) + 
+  labs(x = "Income Percentile", y = "Avg. Death Rate", color = "Indicators") +
+  theme_minimal() +
+  theme(legend.position = "bottom") 
+```
+
+@fig-income_education shows the average death rate across different income percentile and the corresponding average high-education levels.
+The blue line represents the average high-education rates and the red line is the average death rate, corresponding to different y-axis.
+From the graph, it verifies our conclusion which the average death rate decreases as income percentile increases.
+Meanwhile, the proportion of people with at least bachelor degree goes higher.
+
+Therefore, the death rate increases as people voting for the Republican increases.
+However, people living at poorer counties are more likely to vote for Trump and one potential reason behind this is that the the proportion of residents with at least bachelor degree is relatively low in those counties.
+High education level people are more willing to live at the wealthier counties, and they are less likely to vote for Trump.
+Perhaps more educated people do not satisfy the actions taken by the government.
+
+## Income levels to Health insurance: more people can afford the private insurance in wealthier counties and reduce the deaths rate.
+In addition to the political preferences, the healthcare insurance is also significant in predicting the death rate for each county.
+Based on @tbl-second_model, it seems that both `no_insurance` and `private_insurance` are also significant in predicting the death rate for each county but they shows an inverse relationship on the death rate.
+In addition, the values of these two coefficients are almost same, meaning that the effects on death rate between them can be offset.
+
+```{r}
+#| echo: false
+#| message: false
+#| label: fig-income_insurance
+#| fig-cap: The change of average death rate across different education levels
+# Update the summarized_data to include the new averages
+summarized_data <- data %>%
+  mutate(bin = cut(pctile, breaks = seq(0, 100, by = 5), include.lowest = TRUE, labels = seq(2.5, 97.5, by = 5))) %>%
+  group_by(bin) %>%
+  summarise(
+    avg_death_rate = mean(death_rate, na.rm = TRUE),
+    avg_no_insurance = mean(no_insurance, na.rm = TRUE),
+    avg_private_insurance = mean(private_insurance, na.rm = TRUE)
   )
@@
 ggplot(summarized_data, aes(x = as.numeric(as.character(bin)))) +
-  geom_line(aes(y = avg_death_rate, color = "Average Death Rate"), size = 1) +
-  geom_line(aes(y = avg_education_level/2, color = "Average Education Level"), size = 1) + # Adjust the division for better visualization
-  scale_y_continuous(sec.axis = sec_axis(~.*2, name = "Average Education Level")) + # Adjust the multiplier to reverse the transformation
-  labs(x = "Income Percentile", y = "Average Death Rate", color = "Legend") +
-  theme_minimal()
+  geom_line(aes(y = avg_death_rate, color = "Avg. Death Rate"), size = 1) +
+  geom_line(aes(y = avg_no_insurance/8, color = "Avg. No Insurance Rates"), size = 1) +
+  geom_line(aes(y = avg_private_insurance/8, color = "Avg. Private Insurance Rates"), size = 1) + 
+  scale_y_continuous(sec.axis = sec_axis(~.*8, name = "No/Private Insurance Rates")) + 
+  labs(x = "Income Percentile", y = "Avg. Death Rate", color = "Indicators") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
 ```
 
-Education reflects this
+@fig-income_insurance shows the proportion of people with no/private insurance at each level of income.
+Unsurprisingly, as income percentiles increases, the proportion of people with private insurance increases but people with no insurances decreases.
+This suggest that the insurance can be crucial during the pandemic period.
+People with insurance, especially private insurance may be more likely to have medicine or vaccines, hence the counties with more people having insurance are relatively lower compared to the counties which less people having insurance.
+In addition, the private insurance seems to have more significant effects.
 
-## Health insurance iss also an important factor in predicting the death rate of COVID.
-In addition to the political preferences, the healthcare insurance is also significant in predicting the death rate for each county.
-Based on @tbl-second_model, it seems that both `no_insurance` and `private_insurance` are also significant in predicting the death rate for each county but they shows an inverse relationship on the death rate.
-In addition, the values of these two coefficients are almost same, meaning that the effects on death rate between them can be offset.
+# Conclusion
 
-However, t
+This paper looks back on the past three years of the epidemic utilizing the socio-economic, COVID and election data in US to see how the COVID death rate varies across counties in US.
+By the paper, I found that the county's political preference seems to be affect its death rate.
+Low-income people seems to make "stupid" decision, which they vote for the Republican party which brings them more deaths.
+In contrast, people living in the wealthier counties seems to be more smart, perhaps they are more educated.
 
-## The importantance matrix matches our models
+## Weaknesses
 
-## Weaknesses and next steps
+This paper has several limitations.
+First of all, the data of COVID and election are lost for certain counties, especially the counties in Utah.
+In addition, I assumes the models do not have any violations on its assumptions.
+Both drawback may result in inaccurate predictions and lead to the bad performance of the model.
 
-Weaknesses and next steps should also be included.
+## Future research
+For the future research, one primary thing to do is to use a more complicated model such as Multilevel with Post-Stratification (MR.P).
+In addition, it would also be necessary to delve deeper into other factors that may affect the COVID death rate such as the healthcare capacity in each county.
 
@@
 ## Model details
-
-aaa
-
-ojnoj
-
-klmko
-
-j jnjn
 

### 2023-11-01 outputs/paper/paper.qmd
Update the introduction
Second draft

@@
 
-On May 23rd, the WHO chief Tedros announced that the COVID-19 ends and COVID-19 is no longer regarded as a global threat.
-At this point, the epidemic that lasted for three years has officially ended.
-During the three years, nearly 7.6 billion people got infected and about 6.9 million people lost their lives.
-With that said, the US seems to be one of the countries that had influenced by COVID-19 most, there was about 1.1 million deaths and 104 million infections.
-The mortality rate of COVID-19 in US is about 341 per 100,000, which is significantly higher than other western developed countries.
-Meanwhile, the COVID was entirely occured in the Biden's term and next year will be the Federal Election.
-It is worth to know how the COVID deaths is related to the politics and how to recover it, especially that we are now at the post-pandemic period.
+On May 23rd, the WHO's Director-General, Dr.
+Tedros Adhanom Ghebreyesus, announced the end of the COVID-19 pandemic, declaring that the virus was no longer considered a global threat.
+This marked the official end of the pandemic that spanned three years.
+Over this period, nearly 7.6 billion people got infected, and approximately 6.9 million succumbed to the disease.
+The US seems to be one of the countries that have been influenced by COVID-19 the most, with about 1.1 million dead and 104 million infections.
+The mortality rate of COVID-19 in the US is about 3.41 per 1,000, which is significantly higher than in other western developed countries.
+Interestingly, the pandemic began around the time of the 2020 Federal Election and unfolded entirely during President Biden's term.
+As another Federal Election looms next year, it is worth knowing how the COVID deaths are related to socio-economic indicators.
+There's also speculation about the role of political preferences.
+Analyzing these connections could provide insights to better prepare for and prevent future health crises, especially as we navigate the post-pandemic period.
 
-In this paper, I aim to find the variations of death rate across different counties, to see how the counties with different socio-economic factors were affected by COVID differently.
-In addition, I will also include the political factors, to see whether the party that counties voted for will display any differences on COVID death.
-There are three primary data sources which are American Community Survey collecting the socio-economic information for each county, the John-Hopkins public data for COVID cases and deaths for each county and the 2020 US Federal Election from Fox News.
-With capturing this variations, it can help us to access the inequality of mortality caused by COVID across different counties.
-Besides, people living at each county will also know whether their political reference will affect the death rate, this can significantly judge them about how they will vote on the 2024 Federal Election.
+To understand and explore the COVID death rate, I will use five data sets from three different sources.
+American Community Survey [citation] collects the socio-economic information for each county, such as the mean household income.
+The Johns Hopkins University's public dataset [citation] presents data on COVID-19 cases and deaths by county.
+Regarding electoral data, Fox News and The New York Times have comprehensive records of the 2020 US Federal Election, providing the vote counts for each party across all counties.
+These data can help us to assess the inequality of mortality caused by COVID across different countries.
 
-There will be four main parts in this paper: Data, Models, Results, Discussion and Conclusion.
-In the Data part, I will introduce the data used in this paper and high light the key variables.
-The Models and Results session will introduce the model that will be used in this paper and the results from it.
-Moreover, the results will be interpreted and provide the insights about COVID deaths rate.
-Lastly, I will conclude the entire paper and discussion the limitations and drawbacks.
+My analysis will be focused on the variations of death rates across different counties to see how the counties with different socio-economic factors will affect the COVID-19 deaths.
+Moreover, I will incorporate a political dimension, especifically the voting rate in support of the Republican party in each county, to see whether a higher proportion of Republican voters in a county correlates with the COVID-19 death rate.
+Through this paper, I aim to construct a model where the COVID-19 death rate is the dependent variable and includes only those predictors that are statistically significant.
+This study can help US citizens understand the factors affecting the death rate and inform their voting decisions in 2024.
+
+There will be four main parts in this paper.
+In the Data part, I will introduce the data used in this paper and highlight the key variables.
+The Methods and Results parts will introduce the methods used in this paper and the corresponding results.
+Moreover, the results will be interpreted and provide insights about the COVID-19 death rate.
+Lastly, I will conclude the entire paper and discuss the limitations and drawbacks.
 

### 2023-11-01 outputs/paper/paper.qmd
Updating the data part
Second draft

@@
 author: "Yiliu Cao"
-thanks: "Code and data from this analysis are available at: https://github.com/yiliuc/Bicycle_Thefts_2023.git"
+thanks: "Code and data from this analysis are available at: https://github.com/yiliuc/COVID_US_county.git"
 date: "today"
@@
 American Community Survey [citation] collects the socio-economic information for each county, such as the mean household income.
-The Johns Hopkins University's public dataset [citation] presents data on COVID-19 cases and deaths by county.
-Regarding electoral data, Fox News and The New York Times have comprehensive records of the 2020 US Federal Election, providing the vote counts for each party across all counties.
+The Johns Hopkins University's public data set [citation] presents data on COVID-19 cases and deaths by county.
+Regarding electoral data, Fox News have comprehensive records of the 2020 US Federal Election, providing the vote counts for each party across all counties.
 These data can help us to assess the inequality of mortality caused by COVID across different countries.
@@
 
-There will be four main parts in this paper.
-In the Data part, I will introduce the data used in this paper and highlight the key variables.
-The Methods and Results parts will introduce the methods used in this paper and the corresponding results.
-Moreover, the results will be interpreted and provide insights about the COVID-19 death rate.
+There will be five main parts in this paper.
+In the [Data] part, I will introduce the data used in this paper and highlight the key variables.
+The [Methods] and [Results] parts will introduce the methods used in this paper and the corresponding results.
+Moreover, the results will be interpreted and provide insights about the COVID-19 death rate in the [Discussion] part.
 Lastly, I will conclude the entire paper and discuss the limitations and drawbacks.
@@
 
-In this paper, there are three main sources of data, which are American Community Survey (ACS), The Center for Systems Science and Engineering (CSSE) at John-Hopkins University, Fox News.
-These three sources contains different dimensional data in predicting the death rate of COVID in each county.
+As introduced in the introduction, there are three primary sources of data in the data: the American Community Survey (ACS), The Center for Systems Science and Engineering (CSSE) at John Hopkins University [citation], and Fox News [citation].
+Each source provides distinct data dimensions that contribute to predicting the COVID-19 death rate in each US county.
 
-## Source of Data
-
-This paper will use five data sets from three different sources. They are shown below.
-
-### American Community Survey (ACS)
+## American Community Survey (ACS)
 
 The data from ACS contains the socio-economic information for each county.
-While there are many tables from ACS, I will only take three of them which are DP02, DP03 and DP05.
-All the tables are the 2021 five-year ACS estimates
+While there are many tables from ACS, the three tables I will use in this paper are DP02, DP03, and DP05.
+All the tables are the 2021 five-year ACS estimates.
 
-For DP02, it contains the data regarding the social characteristics.
-In this paper, I will only take the information regarding the educational attainment such as the percentage of residents in each county that have at a least bachelor degree.
-Compared to DP02, DP03 contains the information about the economic characteristics.
-To understand the deaths rate in terms of economy, I will grab the information about the proportion of residents with a private insurance, the mean household income and unemployment rate at each county.
-Finally, DP05 contains the demographic data and housing estimates.
-In this paper, I will take the total population, proportions of children and people above 85 and percentage of white and black residents.
+DP02 contains data regarding social characteristics.
+In this paper, I will extract information related to educational attainment, specifically focusing on the percentage of residents in each county who have attained at least a bachelor's degree.
+In contrast, DP03 contains information about the economic characteristics.
+I will take the information about the proportion of residents with private insurance, the mean household income, and the unemployment rate in each county.
+Lastly, DP05 contains the demographic data and housing estimates.
+In this paper, I will take the total population, the proportions of children and individuals above 85, and the percentages of white and black residents in each county.
+The @tbl-DP summarizes the important variables.
 
@@
 #| message: false
+#| label: tbl-DP
 #| tbl-cap: The summary of important variabls from ACS
@@
 
-### CSSE at JHU
+## Center for Systems Science and Engineering at JHU
 
-The CSSE at JHU collects the worldwide COVID data since 2020 to 2023.
-All the raw data are available on their CSSE GitHub which can easily access.
-In particular, the COVID daily report since 2020 for each county on US are listed.
-However, they stop to update the data after March 9th, 2023 as no significant changes.
-The COVID data for US used in the paper is data collected on March 9th.
+The CSSE at JHU has been aggregating global COVID-19 data from 2020 to 2023 for each country in the world.
+All the raw data is readily accessible on their CSSE GitHub repository.
+Notably, they have provided daily COVID-19 reports for each US county since 2020.
+However, they stopped updating the data after March 9th, 2023, as there were no significant changes.
+The COVID data for the US used in the paper was collected on March 9th.
 
-In the data on March 9th, it contains the number of confirmed cases and deaths for each county in US, as well as the incident rate.
-To perform the data analysis later in this paper, I will only use the COVID cases and deaths in each county.
+The data on March 9th contains the number of confirmed cases and deaths for each county in the US, as well as the incident rate.
+For the subsequent data analyses, I will only use the COVID cases and deaths in each county.
+The @tbl-covid summarized them.
 
@@
 #| message: false
+#| label: tbl-covid
 #| tbl-cap: The summary of infection and death rate of COVID in US
@@
 
-Descriptions here
-
 ### Fox News
 
-It contains the county-level election results during the 2020 Federal Election. For each county, it contains the number of votes for the demographic and republican party.
+Fox News has posted the 2020 Election data, which was collected by Tony McGovern [citation].
+It contains the county-level election results during the 2020 Federal Election.
+For each county, it includes the number of votes for the demographic and republican party, as well as the total votes.
 
@@
 
-All data was collected and cleaned using the statistical programming language `R` \[citation\].
-The packages used are `tidyverse` \[citation\], `dplyr` \[citation\], `stringr` \[citation\] and `janitor` \[citation\].
-Besides, `here` \[citation\], `modelsummary` \[citation\], `girdExtra` \[citation\], `usmap` \[citation\], `knitr` \[citation\], `kableExtra` \[citation\] are used to analyse the data.
+All data was collected and cleaned using the statistical programming language `R` [citation].
+The packages used are `tidyverse` [citation], `dplyr` [citation], `stringr` [citation], and `janitor` [citation].
+Besides, `here` [citation], `modelsummary` [citation], `girdExtra` [citation], `usmap` [citation], `knitr` [citation], `kableExtra` [citation] are used to analyze the data.
 
-The procedure to clean the three sources of data are different.
-For all the ACS data downloaded using ACS API \[citation\], referring to the codebook published by ACS \[citation\], I only extract the necessary variables.
-Besides, I clean the name for each county and convert the state name to its abbreviation form.
-In terms of the COVID data from JHU, I only select the cases and deaths and filter only the counties in US.
-Lastly for the election data, I extract the county name and also transform the state name to its abbreviation form.
-I also calculate the number of people voting for Republican party per 1000 voters.
+The way to clean each data set is different.
+For all the ACS data, I downloaded them using ACS API [citation] and only extracted the necessary variables using the codebook published by ACS [citation].
+Additionally, I standardized the county names and transformed state names into their abbreviated forms.
+Regarding the COVID data sourced from JHU, I only select the cases and death numbers of US counties.
+Lastly, for the election data, I not only extracted the county name and transformed the state name to its abbreviation form but also calculated the number of people voting for the Republican party per 1000 voters.
 
-After cleaned each data set, I merge the the five data sets from the above three data sources by the county and state.
+After cleaning each data set, I merge the five data sets by county and state.
 Based on the merged data set, I calculate the infection and death number per 1000 residents in each county.
-Therefore, each row in the merged data represent a US county and provides the the demographic information such as the total population.
-In addition, the data also contains the voting pattern in each county such as the number of votes for democratic and republican party.
+Therefore, each row in the merged data represents a US county and provides demographic information such as the total population.
+In addition, the data also contains the voting pattern in each county.
+The @tbl-summary_variables summarizes all the important variables in this paper.
 
@@
 
-The @tbl-summary_variables summarizes all the important variables in this paper.
-It contains the social characteristics such as high education attainment and economic factors such as income percentile.
-Moreover, the demographic information and political preference for each county are also included
-
 ```{r}
@@
 
-To understand the data better, @fig-boxplot shows the the distribution of death rate for the counties voting for the two parties.
-We can observe that difference in political preference results differences in death rate.
-Counties voting for republican has a relative higher death rate than the counties voting for the Democratic party.
-In addition, there have outliers for both groups.
+To understand the data clearly, @fig-boxplot presents the distribution of death rates for counties based on their voting preference for the two major parties.
+It seems that political preference corresponds with variations in death rates.
+Specifically, counties that leaned Republican exhibited a relatively higher death rate compared to those favoring the Democratic party.
+Notably, outliers are present in both groups.
 
@@
 
-To understand how the death rate correlates with the political preference, @fig-map compares the support for Republican with corresponding death rate.
-From the graph, we can observe that the dark color corresponds to the dark blue, meaning that the counties voting for Republican seem to have a higher death rate than the counties voting for Democratic party.
+To better understand the distribution of death rate from a geography perspective, @fig-map compares the support for Republicans with the corresponding death rate by plotting them onto maps.
+From the graph, we can observe that the dark color corresponds to the dark blue, meaning that the counties voting for the Republican seem to have a higher death rate than those voting for the Democratic Party.
 This pattern is especially significant for the central states.
@@
 
-Our data combines five data sets from three sources. However, it still have limitations which may affect the analyses, reducing the accuracy of our predictions.
+Our data combines five data sets from three sources. Nonetheless, there are still limitations that could influence our results, potentially reducing the accuracy of our predictions.
 
-First of all, there is lack of COVID information in certain states, especially for the state "Utah." We can observe this from @fig-map which some counties shows a color of grey.
-Due to this lackness, the merged data set will also lack the information on these counties even though other data sets have that information.
-The accuracy of the models may be also affected.
-In addition, the data we have may lack of spread, meaning that there could have other information which may also affect the death rate but are not included in this paper such as the hospital capacity.
-Our analyse may be less convincing.
+First, there is a lack of COVID information in certain states, especially in "Utah." This can be observed from @fig-map, where the counties are grey, indicating the data for this county is missing.
+Due to this lack of one data set, the merged data set will also lack information on these counties.
+The accuracy of the subsequent analyses may also be decreased.
 
@@
 https://wisevoter.com/state-rankings/poorest-states/
+
+https://github.com/tonmcg/US_County_Level_Election_Results_08-20

### 2023-11-01 outputs/paper/paper.qmd
Update the methods part
Second draft

@@
 
-As I introduced in the introduction, this paper aim to predict the death rate of COVID in each county of US and to detect whether this correlation related to their political preferences.
-Based on the merged data we have above, I will perform a series of models, from simple to complicated, to detect which predictors are significant in predicting the death rate.
+As I introduced, this paper aims to predict the death rate of COVID in each country of the US and to detect whether this correlation is related to their political preferences.
+Based on the merged data above, I will perform a series of models, from simple to complicated, to detect which predictors are significant in predicting the death rate.
 
@@
 Before we fit the models, I will split the data into training and testing data.
-The training data is used to fit all the models and they are tested on the testing model, to see which one performs betters.
+The training data is used to fit all the models, and they are tested on the testing model to see which one performs better.
 
-For each OLS regression model, the response variable will always be the death rate of COVID in each county.
-However, since the deaths rate can be attributed by various perspectives such as demographic and economics, and we also want to see whether it depends on the party they voted during the last Federal Election.
-I will create two sets of models, which the first set does not consider the political preferences but the second does.
-In addition, for each sets of model, there will be three models with each one considering only one perspective of predictors, an overall model containing all predictors and one best model based on $R^2_{adj}$ and $RMSE$.
-That said, I will fit 10 models in total along with two best models.
-To decide which model is the best one, I will use test them using the testing data, to see which one has a lower testing $RMSE$.
+In each OLS regression model, the dependent variable will consistently be the COVID-19 death rate for each county.
+Given that death rates can be influenced by many factors, including demographics and economics, and given our interest in the potential impact of political party affiliation from the last Federal Election, I will establish two sets of models.
+The first set will not incorporate political preferences, while the second set will.
+Within each of these sets, there will be three models, each focusing exclusively on one category of predictors.
+Additionally, there will be a comprehensive model encompassing all predictors and a best-fit model determined by $R^2_{adj}$ and $RMSE$.
+In total, this approach will yield ten models, including the two best-fit models.
+To decide which model is the best, I will evaluate them using testing data to see which one has a lower testing $RMSE$.
 
-In addition to OLS regression models, I will also draw the regression tree and
-
-The general form of the regression models is: $$Y=\beta_0+\beta_1 X_1+\cdots+\beta_p X_p+\epsilon$$ where: - $Y$ represents the response variables which is the death rate.
+The general form of the regression models is 
+$$Y=\beta_0+\beta_1 X_1+\cdots+\beta_p X_p+\epsilon$$
+where - $Y$ represents the response variables, which is the death rate.
 - $\beta_0$ represents the intercept, which is the death rate holding all variables to zero - $\beta_j$ represents the change of death rates with one unit increase in $X_j$ - $\epsilon$ captures measurement errors and other discrepancies
@@
 After we fit the linear regression models, it is essential to check the assumptions to ensure the accuracy of our predictions.
-There are four main assumptions which are linearity, uncorrelated error, constant variance and normality.
-For the first three assumptions, we can check it by plot the residuals with fitted values.
-However, we need to plot the Normal QQ plot to check the normality.
+There are four main assumptions, which are linearity, uncorrelated error, constant variance, and normality.
+In this paper, I will assume the models satisfy both of the four assumptions.
 
-## Regression Tree and Importance Matrix
+## Expected MSE and Bias-Variance Trade-off
 
-A regression tree is a type of decision tree used specifically for regression problems, where the goal is to predict a continuous outcome variable based on one or more predictor variables.
-By using the regression tress, we can explain the decisions, identify possible events that might occur, and see potential outcomes
+The rationale for partitioning the data into training and testing sets is to validate our model.
+That said, by training a model using training data and then evaluating it using test data, we can determine which model produces the lowest expected MSE and better performance.
 
-The objection function of regression tree is:
+$$
+\mathbb{E}\left[(Y-\hat{f}(x))^2 \mid X=x\right]=\operatorname{Var}(\hat{f}(x))+(\mathbb{E}[\hat{f}(x)]-f(x))^2+\sigma^2
+$$
+The above equation shows the expected MSE, which underscores the bias-variance trade-off.
+It tells us the trade-off between the model's complexity and the interpretability.
+An intricate model may increase the bias, but its variance will be higher if we vary the data, and vice-versa.
 
-$$\min_{j,s} \left[  \sum_{i: x_{i,j} \leq s, x_i \in R1}
-    (y_i - \hat{y}_{R1})^2 +  \sum_{i: x_{i,j} > s, x_i \in R2}
-    (y_i - \hat{y}_{R2})^2 \right]$$
-
-The objective function is to minimize the squared error.
-The $j, s$ represents the index of the variable and threshold, respectively.
-In my regression tree, the two regions, R_1 and R_2, are split by the percentage of the White population in 2020, and the threshold here is 0.094.
-This means all the rows with the percentage of White people in 2020 less or equal to 0.094 are assigned to R1; otherwise, the rest are assigned to R2.
-In addition, the $\hat{y}_{R1}$ and $\hat{y}_{R1}$ represent the mean value of the y in each region.
-According to the regression tree, we can observe that the mean change of percentage vote for Trump is 0.157 and 0.015, which are the values of $\hat{y}_{R1}$ and $\hat{y}_{R2}$.
-
-In addition to regression trees, we can also use the importance matrix which can tell us which predictors are important in predicting the death rate for each county.
-It can help us to verify our final model.
+Hence, by evaluating our models using the testing data, we can know whether we should fit a relatively simple or complicated model.
 

### 2023-11-01 outputs/paper/paper.qmd
Update the Results part
Second draft

@@
 
-## Death rate without considering the political preferences
+## Model Excluding Political Preferences
 
@@
 ```
-Based on the above model, holding all variables to be zero, the expected deaths rate is about 7.2 per 1000.
-Holding other variables fixed, with one percent in the proportion of people who has at least bachelor degree, the death rate is expected to increase by 0.05.
-In addition, with income percentile increases by one, the death rate is expected to decrease by 0.02.
-Unsurprisingly, there is a negative relationship between the proportion of people with no insurance and death rate, 0.7 more people might die if the ratio of people living in a county without insurance increase by 10%.
-In terms of demographic factors, the number of death is inversely correlated with the number of males in a county.
-One percent increase in the males will result 0.06 less death.
-Moreover, it seems that old people are more likely to be died by COVID.
-That said, 0.29 more death if one percent increase in the proportion of people aged above 85.
-Lastly, one percent increase in the proportion of white and black people are both expected to have 0.01 more death.
 
-## Death rate with considering the plotical preference
+The above equation shows the model excluding the political preferences.
+When all variables are set to zero, the anticipated death rate stands at approximately 7.2 per 1,000.
+Keeping other variables constant, an increase of one percent in the proportion of individuals having at least a bachelor's degree is associated with an expected rise of 0.05 in the death rate.
+Furthermore, for every percentile increase in income, the death rate is projected to drop by 0.02.
+Notably, there's a negative correlation between the proportion of uninsured individuals and the death rate; an uptick of 10% in the ratio of uninsured residents in a county may lead to an additional 0.7 deaths per 1,000.
+
+Regarding demographic factors, the number of deaths is inversely correlated with the number of males in a county.
+One percent increase in the males will result in 0.06 less death.
+Moreover, it seems that older people are more likely to die from COVID.
+That said, there will be 0.29 more deaths if there is a one percent increase in the proportion of people aged above 85.
+Lastly, one percent increase in the proportion of white and black people is expected to have 0.01 more deaths.
+
+## Model Including Plotical Preference
 
@@
 ```
-The difference between the second and the first model is that the second model includes the political preference for each county as one predictor to predict the death rate due to COVID.
-Holding other variables, we can see that every 100 more votes per 1000 for republican party, 0.3 more people are expected to dead.
+
+The difference between the second and the first model is that the second model includes the political preference for each county as one predictor to predict the death rate due to COVID-19.
+Holding other variables, we can see that for every 100 more votes per 1000 for the Republican party, 0.3 more people are expected to die.
 
@@
 Comparing the training RMSE from @tbl-all_models_1 and @tbl-all_models_2, we can find that the model containing the political preference has a lower training RMSE.
-However, in order to test the prediction performance of the models we have varying data, I will use testing data to find the test error for the two models above.
+As introduced in the Methods part, to test the prediction performance of the models, I will calculate the test error for the two models to see which is better.
 
@@
 We can observe that the errors for the two models are 1.769733 and 1.675969.
-Therefore, the model containing the political preferences has a lower expecter MSE, hence it has better predictions than another one.
+Therefore, the model containing the political preferences has a lower testing MSE; it has better predictions than the another one.
 

### 2023-11-01 outputs/paper/paper.qmd
Update the citation for R packages

@@
 
-All data was collected and cleaned using the statistical programming language `R` [citation].
-The packages used are `tidyverse` [citation], `dplyr` [citation], `stringr` [citation], and `janitor` [citation].
-Besides, `here` [citation], `modelsummary` [citation], `girdExtra` [citation], `usmap` [citation], `knitr` [citation], `kableExtra` [citation] are used to analyze the data.
+All data was collected and cleaned using the statistical programming language `R` [@R].
+The packages used are `tidyverse` [@tidyverse], `dplyr` [@dplyr], `stringr` [@stringr], and `janitor` [@janitor].
+Besides, `here` [@here], `modelsummary` [@modelsummary], `girdExtra` [@gridExtra], `usmap` [@usmap], `knitr` [@knitr], `kableExtra` [@kableExtra] are used to analyze the data.
 

### 2023-11-01 outputs/paper/paper.qmd
Update the title and abstract

@@
 ---
-title: "aaaa"
+title: "Accessing the COVID-19 Death Rate in the US: While It Varies Across Different Income Levels, the Impact of Political Preference is Unignorable"
 title-block-banner: true
@@
 date-format: "long"
-abstract: "sssss"
+abstract: "This study utilizes the data on socio-economic factors, COVID-19 statistics, and election results sourced from the American Community Survey, the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University, and Fox News.
+Primary findings indicate that the death rate increases as the voting for the Republican party rises, and the wealthier the county, the less the death number.
+While the Democratic Party favors the more affluent counties, the potential reason behind this is that people living in wealthier counties are more educated.
+Moreover, the prevalence of private insurance in these counties could be another factor contributing to fewer deaths.
+Future research should focus on more correlated variables, such as hospital capacity."
 format: pdf

### 2023-11-01 outputs/paper/paper.qmd
Update the reference

@@
 To understand and explore the COVID death rate, I will use five data sets from three different sources.
-American Community Survey [citation] collects the socio-economic information for each county, such as the mean household income.
-The Johns Hopkins University's public data set [citation] presents data on COVID-19 cases and deaths by county.
+American Community Survey [@acs] collects the socio-economic information for each county, such as the mean household income.
+The Center for Systems Science and Engineering (CSSE) at John Hopkins University [@jhu] presents data on COVID-19 cases and deaths by county.
 Regarding electoral data, Fox News have comprehensive records of the 2020 US Federal Election, providing the vote counts for each party across all counties.
@@
 My analysis will be focused on the variations of death rates across different counties to see how the counties with different socio-economic factors will affect the COVID-19 deaths.
-Moreover, I will incorporate a political dimension, especifically the voting rate in support of the Republican party in each county, to see whether a higher proportion of Republican voters in a county correlates with the COVID-19 death rate.
+Moreover, I will incorporate a political dimension, especially the voting rate in support of the Republican party in each county, to see whether a higher proportion of Republican voters in a county correlates with the COVID-19 death rate.
 Through this paper, I aim to construct a model where the COVID-19 death rate is the dependent variable and includes only those predictors that are statistically significant.
@@
 
-As introduced in the introduction, there are three primary sources of data in the data: the American Community Survey (ACS), The Center for Systems Science and Engineering (CSSE) at John Hopkins University [citation], and Fox News [citation].
+As introduced in the introduction, there are three primary sources of data in the data: the American Community Survey (ACS) [@acs], The Center for Systems Science and Engineering (CSSE) at John Hopkins University [@jhu], and Fox News [@fox].
 Each source provides distinct data dimensions that contribute to predicting the COVID-19 death rate in each US county.
@@
 
-Fox News has posted the 2020 Election data, which was collected by Tony McGovern [citation].
+Fox News has posted the 2020 Election data, which was collected by Tony McGovern [@mcgovern].
 It contains the county-level election results during the 2020 Federal Election.
@@
 The way to clean each data set is different.
-For all the ACS data, I downloaded them using ACS API [citation] and only extracted the necessary variables using the codebook published by ACS [citation].
+For all the ACS data, I downloaded them using ACS API [@acs_api] and only extracted the necessary variables using the codebook published by ACS [@acs_codebook].
 Additionally, I standardized the county names and transformed state names into their abbreviated forms.
@@
 plot_grid(plot1, plot2, ncol = 1, align = "v", rel_heights = c(1,1))
+```
+
+```{r}
+#| include: false
+library(ggplot2)
+library(maps)
+states <- map_data("state")
+counties <- map_data("county")
+state_map <- data.frame(state.name = tolower(state.name), state.abb = state.abb)
+counties$region <- state_map$state.abb[match(counties$region, state_map$state.name)]
+counties$subregion <- tools::toTitleCase(counties$subregion)
+
+mer <- merge(counties, data, by.x=c("subregion", "region"), by.y=c("county", "state"), all.x = TRUE)
+mer$death_rate[is.na(mer$death_rate)] <- 0
+
+ggplot(data=mer, mapping=aes(x=long, y=lat, group=group, fill=death_rate)) + 
+  geom_polygon(color="black") +
+  scale_fill_gradient(low="white", high="blue", name="Income") + 
+  ggtitle('U.S. Map with Income by County') + 
+  theme_minimal() + 
+  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
+        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
+  coord_fixed(1.3)
+
+ggplot(mer, aes(x=long, y=lat, group=group)) + 
+  geom_polygon(color="black", fill="gray") +
+  coord_map() +
+  theme_minimal()
+
+duplicated_rows <- data[duplicated(data[c("county", "state")]),]
+if (nrow(duplicated_rows) > 0) {
+  print("There are duplicated rows!")
+  print(duplicated_rows)
+}
 ```

### 2023-11-02 outputs/paper/paper.qmd
Update the Discussion
Second Draft

@@
 library(purrr)
+library(ggplot2)
 
@@
 The packages used are `tidyverse` [@tidyverse], `dplyr` [@dplyr], `stringr` [@stringr], and `janitor` [@janitor].
-Besides, `here` [@here], `modelsummary` [@modelsummary], `girdExtra` [@gridExtra], `usmap` [@usmap], `knitr` [@knitr], `kableExtra` [@kableExtra] are used to analyze the data.
+Besides, `here` [@here], `modelsummary` [@modelsummary], `girdExtra` [@gridExtra], `usmap` [@usmap], `ggpot2` [@ggplot2], `knitr` [@knitr], `kableExtra` [@kableExtra] are used to analyze the data.
 
@@
 plot_grid(plot1, plot2, ncol = 1, align = "v", rel_heights = c(1,1))
-```
-
-```{r}
-#| include: false
-library(ggplot2)
-library(maps)
-states <- map_data("state")
-counties <- map_data("county")
-state_map <- data.frame(state.name = tolower(state.name), state.abb = state.abb)
-counties$region <- state_map$state.abb[match(counties$region, state_map$state.name)]
-counties$subregion <- tools::toTitleCase(counties$subregion)
-
-mer <- merge(counties, data, by.x=c("subregion", "region"), by.y=c("county", "state"), all.x = TRUE)
-mer$death_rate[is.na(mer$death_rate)] <- 0
-
-ggplot(data=mer, mapping=aes(x=long, y=lat, group=group, fill=death_rate)) + 
-  geom_polygon(color="black") +
-  scale_fill_gradient(low="white", high="blue", name="Income") + 
-  ggtitle('U.S. Map with Income by County') + 
-  theme_minimal() + 
-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
-        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
-  coord_fixed(1.3)
-
-ggplot(mer, aes(x=long, y=lat, group=group)) + 
-  geom_polygon(color="black", fill="gray") +
-  coord_map() +
-  theme_minimal()
-
-duplicated_rows <- data[duplicated(data[c("county", "state")]),]
-if (nrow(duplicated_rows) > 0) {
-  print("There are duplicated rows!")
-  print(duplicated_rows)
-}
 ```
@@
 
-## Politic preference to income levels: wealthier people are less likely to vote for the Republican, perhaps they are more educated.
+## Although income levels are known to affect the COVID mortality, the impact of political preferences is more substantial than anticipated, potentially overcshadowing the effects of income.
 
@@
 
+```{r, fig.width=10, fig.height=4}
+#| echo: false
+#| message: false
+#| label: fig-income_state
+#| fig-cap: Variation in COVID-19 Mortality Rates Across States, Categorized by Income Levels and Dominant Political Preferences
+
+library(dplyr)
+library(ggplot2)
+# Find the wining data for each state
+data_win_party <- data %>%
+  group_by(state) %>%
+  summarise(dem_votes = sum(votes_dem),
+            rep_votes = sum(votes_rep), 
+            .groups = 'drop') %>% 
+  mutate(winning_party = ifelse(dem_votes > rep_votes, 'Democratic', 'Republican')) %>% 
+  select(state, winning_party)
+# Calculate the average income
+average_income_by_state <- data %>%
+  group_by(state) %>%
+  summarise(average_income = mean(mean_household_income), .groups = 'drop')
+# Join the winning party and average income back to the original data
+data_win_party_income <- data %>%
+  left_join(data_win_party, by = "state") %>%
+  left_join(average_income_by_state, by = "state")
+# Arrange states by average income
+ordered_states <- data_win_party_income %>%
+  arrange(average_income) %>%
+  .$state %>%
+  unique()
+# Create a factor with levels based on income-ordered states for plotting
+data_win_party_income$state <- factor(data_win_party_income$state, levels = ordered_states)
+
+average_death_rate <- mean(data_win_party_income$death_rate)
+
+ggplot(data_win_party_income, aes(x = state, y = death_rate, fill = winning_party)) +
+  geom_boxplot() +
+  geom_hline(yintercept = average_death_rate, linetype = "dashed", color = "black") +
+  annotate("text", x = Inf, y = average_death_rate, label = " National Average", hjust = 1, vjust = -0.5, color = "black", size = 3) +
+
+  # Adding the dashed line
+  scale_fill_manual(values = c('Democratic' = 'blue', 'Republican' = 'red')) +
+  labs(y = 'Death Rate', x = 'State (Ascending Income)', fill = "Winning Party") +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+```
+
+@fig-income_state can prove our point.
+The @fig-income_state shows the spread of COVID death rate for each state in the order of ascend income levels.
+The color represent the party each state supported during 2020 US Federal Election.
+We can observe that twelve out of the top 15 poorest counties voted for Donald Trump.
+In contrast, only one of the top 15 richest states voted for the Republican Party.
+In terms of the death rate, the @fig-income_state also shows that the death rate decreases as income level increases and majority of counties voting for Democratic party has a death rate lower than the national average level.
+
+```{r}
+#| echo: false
+#| message: false
+library(dplyr)
+library(ggplot2)
+
+# Assuming 'data' is your DataFrame with columns: pctile, rep_rate, dem_rate
+
+# Step 1: Calculate the average support rate for each party at each income percentile
+avg_support <- data %>%
+  group_by(pctile) %>%
+  summarise(
+    avg_rep_rate = mean(rep_rate),
+    avg_dem_rate = mean(dem_rate)
+  ) %>%
+  ungroup()
+
+# Step 2: Divide income percentile into four groups
+avg_support <- avg_support %>%
+  mutate(
+    pctile_group = cut(pctile, breaks = quantile(pctile, probs = 0:4/4), include.lowest = TRUE)
+  )
+
+# Step 3: Fit regression models and predict values for each piece (this is a placeholder)
+# You would fit your piece-wise regression models here for each party and each group
+
+# Plotting with placeholder regression lines (You'll need the actual fitted model predictions)
+ggplot(avg_support, aes(x = pctile)) +
+  geom_point(aes(y = avg_rep_rate), color = 'red') +
+  geom_point(aes(y = avg_dem_rate), color = 'blue') +
+  # Placeholder for regression lines
+  geom_smooth(data = subset(avg_support, pctile_group == 1), aes(y = avg_rep_rate, color = "Republican"), method = "lm", se = FALSE) +
+  geom_smooth(data = subset(avg_support, pctile_group == 1), aes(y = avg_dem_rate, color = "Democrat"), method = "lm", se = FALSE) +
+  # ... repeat geom_smooth for other groups ...
+  scale_color_manual(values = c("Republican" = "red", "Democrat" = "blue")) +
+  labs(title = 'Average Support Rate by Income Percentile',
+       x = 'Income Percentile', y = 'Average Support Rate') +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+# Note: In a real scenario, you would repeat the geom_smooth for the other pctile_groups (2, 3, 4)
+# and would use the actual fitted model predictions.
+
+```
+
 This is a surprising but also horrible observation as Donald Trump is especially in favor of poor people.
-Eight out of the top 10 counties with highest poverty rate voted for Donald Trump.
-Referring to @fig-map, we can clearly see that the above 80% states in the poorer states, middle US, had voted for Trump.
-Conversely, above 80% wealthier coastal states, such as California and New York, had voted for The Democratic party.
-These both verify our observations from the model.
-What Trump's votes does not know is that their support may bring more deaths to their counties.
+This means the low-income voters voting for Trump increase their chances of death in disguised form.
+Based on the fact that the death of COVID of low-income people is almost two times higher than the high-income [citation], it seems that it is not unfair to attribute the high death rate of US to Republican Party (Donald Trump).
 
-But why, why the counties with higher income levels are less likely to vote for the Republican party and hence results in a lower death rate?
-One thing that can explain this pattern from our model is the proportion of people that have higher education level.
-According to @tbl-second_model, we can find that there is an inverse relationship between the death rate and the number of residents with high education level.
-Its value of coefficient is much higher than the previous two, indicating that education level is an essential factor in predicting the death rate.
+## The difference in high-education atteinment and insurance structure may be potential factors causing the above phenomenon.
+
+The model [@tbl-second_model] incorporates the proportion of people that have at least bachelor degree, and the proportion of people with private insurance or without insurance.
+In terms of education level, there is an inverse relationship between the death rate and the number of residents with high education level.
+Its coefficient value is relatively high, the death number can be decreased by 0.24 if there are ten more percent people having higher degree.
 
@@
 
-Therefore, the death rate increases as people voting for the Republican increases.
-However, people living at poorer counties are more likely to vote for Trump and one potential reason behind this is that the the proportion of residents with at least bachelor degree is relatively low in those counties.
-High education level people are more willing to live at the wealthier counties, and they are less likely to vote for Trump.
-Perhaps more educated people do not satisfy the actions taken by the government.
+Therefore, one reason can explain the inequality of COVID mortality is that there are less people with higher degree lives in the low-income county.
+High educational level people are more willing to live at the wealthier counties.
+They are more educated and perhaps they know what Trump did was wrong and hence they did not vote for Trump.
+This can also explain why there are less votes for Trump in wealthier counties/states.
 
-## Income levels to Health insurance: more people can afford the private insurance in wealthier counties and reduce the deaths rate.
-In addition to the political preferences, the healthcare insurance is also significant in predicting the death rate for each county.
+In addition to education level, the healthcare insurance is also significant in predicting the death rate for each county.
 Based on @tbl-second_model, it seems that both `no_insurance` and `private_insurance` are also significant in predicting the death rate for each county but they shows an inverse relationship on the death rate.
@@
 In addition, the private insurance seems to have more significant effects.
+
+Comparing education with insurance, one common thing is that the we can see diverging trends in poorer and wealthier counties.
+People living in wealthier counties are more educated and concerns more about the public issue.
+Meanwhile, their awareness of protecting themselves from COVID is stronger than the poorer people.
+Due to their wealth level, they have the money to purchase private insurance and are more accessible to the medicines or vaccines.
+This explains why the wealthier counties had a relatively lower COVID mortality rate than poor ones.
 
@@
 
-# References
+# References {.unnumbered}
 
@@
 https://github.com/tonmcg/US_County_Level_Election_Results_08-20
+
+https://www.reuters.com/world/us/us-poor-died-much-higher-rate-covid-than-rich-report-2022-04-04/

### 2023-11-03 scripts/02-data_cleaning.R
Update the cleaned data

@@ -110,7 +110,8 @@ election_data_clean <- election_data_clean %>%
          rep_win = case_when(votes_rep > votes_dem ~ 1,
                              votes_rep < votes_dem ~ 0,
                              TRUE ~ 1),
-         rep_rate = (votes_rep/total_votes)*1000,)
+         rep_rate = (votes_rep/total_votes)*1000,
+         dem_rate = (votes_dem/total_votes)*1000)
 election_data_clean <- na.omit(election_data_clean)
 
 # Merge the five data sets. The merged data set will be the data in this paper
@@ -121,21 +122,6 @@ merged_data_acs_covid <- merged_data_acs %>%
   inner_join(covid_data_clean, by = c("state", "county"))
 merged_data_overall <- merged_data_acs_covid %>% 
   inner_join(election_data_clean, by = c("state", "county"))
-
-
-# unique_rows_DP05 <- DP05_clean %>%
-#   anti_join(covid_data_clean, by = c("state", "county"))
-# 
-# unique_rows_election <- election_data_clean %>%
-#   anti_join(covid_data_clean, by = c("state", "county"))
-# 
-# unique_rows_election <- election_data_clean %>%
-#   anti_join(covid_data_clean, by = c("state", "county"))
-# 
-# unique_rows_covid <- covid_data_clean %>%
-#   anti_join(election_data_clean, by = c("state", "county"))
-
-
 
 merged_data_acs_covid$infection_rate <- (merged_data_acs_covid$cases/merged_data_acs_covid$total_population)*1000
 merged_data_acs_covid$death_rate <- (merged_data_acs_covid$deaths/merged_data_acs_covid$total_population)*1000

### 2023-11-03 outputs/paper/paper.qmd
Done!

@@
 date-format: "long"
-abstract: "This study utilizes the data on socio-economic factors, COVID-19 statistics, and election results sourced from the American Community Survey, the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University, and Fox News.
-Primary findings indicate that the death rate increases as the voting for the Republican party rises, and the wealthier the county, the less the death number.
-While the Democratic Party favors the more affluent counties, the potential reason behind this is that people living in wealthier counties are more educated.
-Moreover, the prevalence of private insurance in these counties could be another factor contributing to fewer deaths.
+abstract: "This study utilizes the data on socio-economic factors, COVID-19 statistics, and election results sourced from the American Community Survey, CSSE at JHU and Fox News.
+Primary findings indicate that the death rate positively relates to voting for the Republican party but has an inverse relation to income levels.
+The research further identifies a trend of Republican party preference among low-income counties, and the counties voting for the Democratic exhibit lower mortality rates.
+Furthermore, the potential reasons contributing to the inequality in mortality rate might be the high education and insurance system.
 Future research should focus on more correlated variables, such as hospital capacity."
@@
 On May 23rd, the WHO's Director-General, Dr.
-Tedros Adhanom Ghebreyesus, announced the end of the COVID-19 pandemic, declaring that the virus was no longer considered a global threat.
+Tedros Adhanom Ghebreyesus, announced the end of the COVID-19 pandemic [@un], declaring that the virus was no longer considered a global threat.
 This marked the official end of the pandemic that spanned three years.
 Over this period, nearly 7.6 billion people got infected, and approximately 6.9 million succumbed to the disease.
-The US seems to be one of the countries that have been influenced by COVID-19 the most, with about 1.1 million dead and 104 million infections.
+The US seems to be one of the countries that have been influenced by COVID-19 the most, with about 1.1 million dead and 104 million infections [@aljazeera_2023].
 The mortality rate of COVID-19 in the US is about 3.41 per 1,000, which is significantly higher than in other western developed countries.
@@
 The above two models are fitted using the training data.
-Comparing the training RMSE from @tbl-all_models_1 and @tbl-all_models_2, we can find that the model containing the political preference has a lower training RMSE.
+Comparing the training RMSE from @tbl-all_models1 and @tbl-all_models2, we can find that the model containing the political preference has a lower training RMSE.
 As introduced in the Methods part, to test the prediction performance of the models, I will calculate the test error for the two models to see which is better.
@@
 
-To sum up, the second model which contains the political reference is more appropriate than another one, representing that the party each county supported during the last election may change the attitude that people face with COVID and hence results differences in death rate.
-In addition to that, it is also important to consider the rest variables while predicting the death rate.
+To sum up, the previous results show that it is essential to consider the political reference when predicting the COVID mortality rate.
+However, combining political preference with other demographic and economic factors will be more convincing and interpretable.
+The key takeaways are as follows.
 
-## Although income levels are known to affect the COVID mortality, the impact of political preferences is more substantial than anticipated, potentially overcshadowing the effects of income.
+## Although income levels are typically associated with variations in COVID-19 mortality, the influence of political preferences appears to be more significant than anticipated.
 
-From the @tbl-second_model, we can observe that both income percentile and the support of Republican both are significant in predicting the death rate of COVID, but with an inverse direction of their coefficients.
-More people voting for Republican party is expected to increase the number of deaths due to COVID in this county and a wealthier counties usually have a less COVID deaths.
-However, if we compare their values of coefficients, we can find that the increase in death by every three more people in 1000 voting for Republican party is equivalent to decrease in death rate by one increase in the income percentile.
-In other words, counties with the lowest income levels might be the worst areas sacrificing from COVID if they support the Republican party.
+From the @tbl-second_model, both income percentile and the support of Republicans are significant in predicting the death rate of COVID-19, but with opposing effects.
+More people voting for the Republican party correlates with a rise in mortality; conversely, higher-income counties are expected to have a lower death rate.
+In terms of the values of coefficients, the coefficient for the income percentile is three times as large as the one for voting for Republican, meaning that for every additional three out of a thousand individuals voting Republican, the increase in death rate is equivalent to the decrease in the death rate by a one-percentile rise in income level.
+This suggests that the lowest-income counties may suffer disproportionately from COVID-19 fatalities, particularly where Republican support is vital.
 
-```{r, fig.width=10, fig.height=4}
+```{r, fig.width=10, fig.height=4.5}
 #| echo: false
@@
   scale_fill_manual(values = c('Democratic' = 'blue', 'Republican' = 'red')) +
-  labs(y = 'Death Rate', x = 'State (Ascending Income)', fill = "Winning Party") +
+  labs(y = 'Death Rate', x = 'State (Ascending Income)', 
+       fill = "Winning Party") +
   theme_minimal() +
@@
 
-@fig-income_state can prove our point.
-The @fig-income_state shows the spread of COVID death rate for each state in the order of ascend income levels.
-The color represent the party each state supported during 2020 US Federal Election.
-We can observe that twelve out of the top 15 poorest counties voted for Donald Trump.
-In contrast, only one of the top 15 richest states voted for the Republican Party.
-In terms of the death rate, the @fig-income_state also shows that the death rate decreases as income level increases and majority of counties voting for Democratic party has a death rate lower than the national average level.
+@fig-income_state corroborates our assertions.
+It shows the spread of the COVID-19 death rate for each state in the order of ascending income levels.
+The colour represents the party each state supported during the 2020 US Federal Election.
+Notably, twelve out of the top 15 poorest counties exhibited a preference for Donald Trump, while, in stark contrast, only one among the fifteen wealthiest states favoured the Republican Party.
+Regarding mortality, the @fig-income_state also shows that the death rate decreases as income level increases, and the majority of counties voting for the Democratic party have a death rate lower than the national average level.
+
+Compared with @fig-income_state, @fig-voters_party examines the number of voters across the income percentile.
+It shows that Republicans are favoured among lower-income groups, while Democratic voters have grown significantly, corresponding to the wealthiest percentile.
 
@@
 #| message: false
+#| label: fig-voters_party
+#| fig-cap: The average votes for the two party at each income percentile
+
 library(dplyr)
 library(ggplot2)
-
-# Assuming 'data' is your DataFrame with columns: pctile, rep_rate, dem_rate
-
-# Step 1: Calculate the average support rate for each party at each income percentile
 avg_support <- data %>%
@@
   summarise(
-    avg_rep_rate = mean(rep_rate),
-    avg_dem_rate = mean(dem_rate)
+    avg_votes_rep = mean(votes_rep/1000),
+    avg_votes_dem = mean(votes_dem/1000)
   ) %>%
   ungroup()
-
-# Step 2: Divide income percentile into four groups
 avg_support <- avg_support %>%
@@
   )
+library(dplyr)
+library(ggplot2)
 
-# Step 3: Fit regression models and predict values for each piece (this is a placeholder)
-# You would fit your piece-wise regression models here for each party and each group
+# Assuming 'avg_support' is your data frame with the calculated averages and percentile groups
 
-# Plotting with placeholder regression lines (You'll need the actual fitted model predictions)
-ggplot(avg_support, aes(x = pctile)) +
-  geom_point(aes(y = avg_rep_rate), color = 'red') +
-  geom_point(aes(y = avg_dem_rate), color = 'blue') +
-  # Placeholder for regression lines
-  geom_smooth(data = subset(avg_support, pctile_group == 1), aes(y = avg_rep_rate, color = "Republican"), method = "lm", se = FALSE) +
-  geom_smooth(data = subset(avg_support, pctile_group == 1), aes(y = avg_dem_rate, color = "Democrat"), method = "lm", se = FALSE) +
-  # ... repeat geom_smooth for other groups ...
-  scale_color_manual(values = c("Republican" = "red", "Democrat" = "blue")) +
-  labs(title = 'Average Support Rate by Income Percentile',
-       x = 'Income Percentile', y = 'Average Support Rate') +
+# Plot the points for each party
+p <- ggplot(avg_support, aes(x = pctile)) +
+  geom_point(aes(y = avg_votes_rep, color = "Republican"), alpha = 0.5) +
+  geom_point(aes(y = avg_votes_dem, color = "Democratic"), alpha = 0.5) +
+  geom_vline(xintercept = c(25, 50, 75), 
+             linetype = "dashed", color = "grey", size = 0.5)
+
+# Add piece-wise regression lines for each group and party
+for(group in unique(avg_support$pctile_group)) {
+  p <- p + geom_smooth(data = avg_support %>% filter(pctile_group == group),
+                       aes(y = avg_votes_rep, group = group),
+                       method = "lm", color = 'red', se = FALSE, formula = y ~ x) +
+    geom_smooth(data = avg_support %>% filter(pctile_group == group),
+                aes(y = avg_votes_dem, group = group),
+                method = "lm", color = 'blue', se = FALSE, formula = y ~ x)
+}
+
+# Finalize the plot with custom settings and labels
+p <- p + scale_color_manual(values = c("Republican" = "red", "Democratic" = "blue")) +
+  labs(x = 'Income Percentile', 
+       y = 'Average Voters (K)',
+       colour = "Political Party") +
   theme_minimal() +
   theme(legend.position = "bottom")
-
-# Note: In a real scenario, you would repeat the geom_smooth for the other pctile_groups (2, 3, 4)
-# and would use the actual fitted model predictions.
-
+print(p)
 ```
 
-This is a surprising but also horrible observation as Donald Trump is especially in favor of poor people.
-This means the low-income voters voting for Trump increase their chances of death in disguised form.
-Based on the fact that the death of COVID of low-income people is almost two times higher than the high-income [citation], it seems that it is not unfair to attribute the high death rate of US to Republican Party (Donald Trump).
+Combining the @fig-income_state and @fig-voters_party, they yield an intriguing yet concerning observation.
+They suggest that lower-income voters, who disproportionately voted for Donald Trump, may inadvertently face higher mortality risks.
+This observation is underpinned by evidence indicating that COVID-19 mortality rates among low-income populations are nearly double those in higher-income brackets [@thomson_reuters_2022].
+While it is critical to approach causality with caution, these findings seem to suggest that Trump has an inescapable responsibility for the high death under his misleading during the pandemic.
 
-## The difference in high-education atteinment and insurance structure may be potential factors causing the above phenomenon.
+## The difference in high-education attainment and insurance structure may be potential factors causing the above phenomenon.
 
-The model [@tbl-second_model] incorporates the proportion of people that have at least bachelor degree, and the proportion of people with private insurance or without insurance.
-In terms of education level, there is an inverse relationship between the death rate and the number of residents with high education level.
-Its coefficient value is relatively high, the death number can be decreased by 0.24 if there are ten more percent people having higher degree.
+The inequality observed in COVID-19 mortality rates between low and high-income areas in the US may be potentially explained by the differences in educational attainment and insurance coverage, as suggested in @tbl-second_model.
+The model shows an inverse correlation between mortality rates and the percentage of individuals with high education levels: for every 10% increase in the population attaining a bachelor's degree or higher, there is a corresponding 0.24 decrease in the number of deaths.
 
@@
 
-@fig-income_education shows the average death rate across different income percentile and the corresponding average high-education levels.
-The blue line represents the average high-education rates and the red line is the average death rate, corresponding to different y-axis.
-From the graph, it verifies our conclusion which the average death rate decreases as income percentile increases.
-Meanwhile, the proportion of people with at least bachelor degree goes higher.
+@fig-income_education shows the average death rate across different income percentiles and the corresponding average proportion of the population attaining a bachelor's degree or higher.
+The blue line represents the average high education rates, and the red line is the average death rate.
+From the graph, as the income levels increase, more people have at least a bachelor's degree and the average death rate decreases.
 
-Therefore, one reason can explain the inequality of COVID mortality is that there are less people with higher degree lives in the low-income county.
-High educational level people are more willing to live at the wealthier counties.
-They are more educated and perhaps they know what Trump did was wrong and hence they did not vote for Trump.
-This can also explain why there are less votes for Trump in wealthier counties/states.
+Therefore, one possible reason explaining the inequality of COVID mortality is that individuals with higher degrees may prefer to live in more affluent areas.
+These populations are generally more informed and may have differing perspectives on the effectiveness of the policies during the Trump administration, which may explain the lower support for Trump in wealthier counties or states.
 
-In addition to education level, the healthcare insurance is also significant in predicting the death rate for each county.
-Based on @tbl-second_model, it seems that both `no_insurance` and `private_insurance` are also significant in predicting the death rate for each county but they shows an inverse relationship on the death rate.
-In addition, the values of these two coefficients are almost same, meaning that the effects on death rate between them can be offset.
+In addition to education level, healthcare insurance is another significant in predicting the death rate for each county, as suggested on @tbl-second_model.
+The variables `no_insurance` and `private_insurance` demonstrate substantial inverse correlations with death rates.
+Interestingly, the magnitudes of their coefficients are nearly identical, suggesting that the presence of private insurance potentially counterbalances the impact on mortality rates from the lack of insurance.
 
@@
   theme_minimal() +
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom") +
+  guides(color = guide_legend(nrow = 2))
 ```
 
-@fig-income_insurance shows the proportion of people with no/private insurance at each level of income.
-Unsurprisingly, as income percentiles increases, the proportion of people with private insurance increases but people with no insurances decreases.
-This suggest that the insurance can be crucial during the pandemic period.
-People with insurance, especially private insurance may be more likely to have medicine or vaccines, hence the counties with more people having insurance are relatively lower compared to the counties which less people having insurance.
-In addition, the private insurance seems to have more significant effects.
+@fig-income_insurance delineates the correlation between income levels and the proportion of individuals with varying types of health insurance.
+Unsurprisingly, as income percentiles increase, more people have private insurance increases, and fewer people have no insurance.
+This pattern underscores the pivotal role of insurance during the pandemic, as people with insurance, especially private insurance,e may be more likely to have medicine or vaccines.
+Hence, the counties with more people having insurance are relatively lower compared to the counties with fewer people having insurance.
 
-Comparing education with insurance, one common thing is that the we can see diverging trends in poorer and wealthier counties.
-People living in wealthier counties are more educated and concerns more about the public issue.
-Meanwhile, their awareness of protecting themselves from COVID is stronger than the poorer people.
-Due to their wealth level, they have the money to purchase private insurance and are more accessible to the medicines or vaccines.
-This explains why the wealthier counties had a relatively lower COVID mortality rate than poor ones.
+When considering both education and insurance, there is a convergent pattern among the diverging trends in poorer and wealthier countries.
+Residents of more affluent counties not only tend to be better educated but also exhibit a higher awareness about their health.
+This is possible because people living in wealthier counties care more about public health issues, and they can also afford the cost of private insurance.
+This composite of socio-economic advantages elucidates the observed discrepancy in COVID-19 mortality rates, with wealthier counties experiencing fewer fatalities than their less affluent counterparts.
 
@@
 
-This paper looks back on the past three years of the epidemic utilizing the socio-economic, COVID and election data in US to see how the COVID death rate varies across counties in US.
-By the paper, I found that the county's political preference seems to be affect its death rate.
-Low-income people seems to make "stupid" decision, which they vote for the Republican party which brings them more deaths.
-In contrast, people living in the wealthier counties seems to be more smart, perhaps they are more educated.
+This paper reflects on the socio-economic dynamics, COVID-19 mortality, and electoral outcomes in the United States over the past three years of the pandemic, examining variations in death rates across counties.
+The primary findings suggest that political preferences may have a more pronounced correlation with COVID-19 death rates than income levels.
+By illustrating the interplay between these two factors, it suggests a trend where lower-income counties, which showed a higher propensity to vote for Donald Trump, also experienced higher mortality rates.
+Throughout the model, education and insurance coverage seem to be the two key factors explaining the inequality of mortality rate between the wealthier and poorer counties.
+More educated people are richer, and they can afford the private insurance cost, meaning that they are more accessible to those healthcare facilities and hence have a lower death rate.
+
+Since the political data used in this paper was the last US Federal Election results, this study could offer insights into the electoral landscape for the forthcoming Federal Election.
+By observing the death rate in the counties or states voting for Joe Biden, we can see that they have a lower death rate than the counties voting for Donald Trump.
+It may suggest that the Democratic party may be able to implement more effective public health measures.
+This can bring insights regarding the next federal election, which Joe Biden may be more likely to reelect.
 
@@
 This paper has several limitations.
-First of all, the data of COVID and election are lost for certain counties, especially the counties in Utah.
-In addition, I assumes the models do not have any violations on its assumptions.
-Both drawback may result in inaccurate predictions and lead to the bad performance of the model.
+First, the data on COVID and elections are lost for certain counties, especially the counties in Utah.
+In addition, I assume the models do not have any violations of their assumptions.
+Besides, the way that I choose the best model is to find the model in which all predictors are significant.
+There should be a more complicated procedure when selecting the models.
+With these drawbacks, it may lead to a bad performance of the models, and hence, the predictions might be inaccurate.
 
 ## Future research
-For the future research, one primary thing to do is to use a more complicated model such as Multilevel with Post-Stratification (MR.P).
-In addition, it would also be necessary to delve deeper into other factors that may affect the COVID death rate such as the healthcare capacity in each county.
+
+For future research, one primary thing to do is to use a more complicated model, such as Multilevel with Post-Stratification (MR.P).
+In addition, it would also be necessary to delve deeper into other factors that may affect the COVID-19 death rate, such as the healthcare capacity in each county.
+Furthermore, it will also be necessary to find the causal inference between the Electrol outcome and COVID.
 
@@
 #| message: false
-#| label: tbl-all_models_1
+#| label: tbl-all_models1
 #| tbl-cap: Summary of the set of models with political preferences
@@
 # References {.unnumbered}
-
-https://news.un.org/en/story/2023/05/1136367 https://www.aljazeera.com/news/2023/5/11/three-years-1-1-million-deaths-covid-emergency-ending-in-us
-https://wisevoter.com/state-rankings/poorest-states/
-
-https://github.com/tonmcg/US_County_Level_Election_Results_08-20
-
-https://www.reuters.com/world/us/us-poor-died-much-higher-rate-covid-than-rich-report-2022-04-04/

### 2023-11-11 outputs/paper/paper.qmd
Done!
Update tiny format error
Revise the paper
Except CV and explainations

@@
 ---
-title: "Accessing the COVID-19 Death Rate in the US: While It Varies Across Different Income Levels, the Impact of Political Preference is Unignorable"
+title: "Accessing the COVID-19 Mortality Rate in the US: While It Varies Across Different Income Levels, the Impact of Political Preference is Unignorable"
 title-block-banner: true
@@
 abstract: "This study utilizes the data on socio-economic factors, COVID-19 statistics, and election results sourced from the American Community Survey, CSSE at JHU and Fox News.
-Primary findings indicate that the death rate positively relates to voting for the Republican party but has an inverse relation to income levels.
+Primary findings indicate that the mortality rate positively relates to voting for the Republican party but has an inverse relation to income levels.
 The research further identifies a trend of Republican party preference among low-income counties, and the counties voting for the Democratic exhibit lower mortality rates.
@@
 
-On May 23rd, the WHO's Director-General, Dr.
+On May 23rd 2023, the WHO's Director-General, Dr.
 Tedros Adhanom Ghebreyesus, announced the end of the COVID-19 pandemic [@un], declaring that the virus was no longer considered a global threat.
 This marked the official end of the pandemic that spanned three years.
-Over this period, nearly 7.6 billion people got infected, and approximately 6.9 million succumbed to the disease.
+Over this period, nearly 7.6 billion confired cases, and approximately 6.9 million succumbed to the disease.
 The US seems to be one of the countries that have been influenced by COVID-19 the most, with about 1.1 million dead and 104 million infections [@aljazeera_2023].
@@
   geom_boxplot() +
+  geom_jitter(alpha = 0.2, width = 0.3, height = 0) +
   theme_minimal() +
@@
 plot1 <- plot_usmap(data = data, values = "rep_rate", regions = "counties", exclude = c("AK", "Hawaii")) + 
-  scale_fill_gradient2(name = "Voting for the Republican %", 
+  scale_fill_gradient2(name = "Voting for the Republican per 1000", 
                        low = "blue", mid = "white", high = "red",
@@
                color = "black", size = 0.8) +  
-  labs(subtitle = "The voting patter for each US county") +
+  labs(subtitle = "The voting pattern for each US county") +
   theme(plot.subtitle = element_text(size = 12,
@@
 
-# Arrange the plots vertically with a main title
 library(cowplot)
@@
 
-To better understand the distribution of death rate from a geography perspective, @fig-map compares the support for Republicans with the corresponding death rate by plotting them onto maps.
+To better understand the distribution of death rate from a geographic perspective, @fig-map compares the support for Republicans with the corresponding death rate by plotting them onto maps.
 From the graph, we can observe that the dark color corresponds to the dark blue, meaning that the counties voting for the Republican seem to have a higher death rate than those voting for the Democratic Party.
@@
 $$Y=\beta_0+\beta_1 X_1+\cdots+\beta_p X_p+\epsilon$$
-where - $Y$ represents the response variables, which is the death rate.
-- $\beta_0$ represents the intercept, which is the death rate holding all variables to zero - $\beta_j$ represents the change of death rates with one unit increase in $X_j$ - $\epsilon$ captures measurement errors and other discrepancies
+where:
+
+- $Y$ represents the response variables, which is the death rate.
+- $\beta_0$ represents the intercept, which is the death rate holding all variables to zero.
+- $\beta_j$ represents the change of death rates with one unit increase in $X_j$.
+- $\epsilon$ captures measurement errors and other discrepancies
 
@@
 
-## Model Including Plotical Preference
+## Model Including Political Preference
 
@@
 The difference between the second and the first model is that the second model includes the political preference for each county as one predictor to predict the death rate due to COVID-19.
-Holding other variables, we can see that for every 100 more votes per 1000 for the Republican party, 0.3 more people are expected to die.
+Holding other variables constant, we can see that for every 100 more votes per 1000 for the Republican party, 0.3 more people are expected to die.
 
@@
 
-## Model details
+## Model details {.unnumbered}
 

### 2023-11-17 outputs/paper/paper.qmd
Revise the paper

@@
 library(ggplot2)
+library(boot)
 
@@
 Each source provides distinct data dimensions that contribute to predicting the COVID-19 death rate in each US county.
+The @tbl-summary_variables introduces all the important variables in this paper.
+
+```{r}
+#| echo: false
+#| message: false
+#| label: tbl-summary_variables
+#| tbl-cap: Summary of important variables in the model
+data.frame(variable = c("High education attainment",
+                        "Income Percentile",
+                        "People with no insurance",
+                        "People with private insurance",
+                        "Males",
+                        "People aged above 85",
+                        "White people",
+                        "Black people",
+                        "People supporting Republican"),
+           coded_name = c("high_education",
+                          "IncomePctile",
+                          "no_insurance",
+                          "private_insurance",
+                          "males",
+                          "age_85",
+                          "white",
+                          "black",
+                          "rep_rate"),
+           description = c("The proportion of local residents with at least Bachelor degree",
+                           "The mean household income percentile representing the income level for each county",
+                           "The proportion of local residents without insurance",
+                           "The proportion of local residents with private insurance",
+                           "The proportion of male residents",
+                           "The proportion of people aged above 85",
+                           "The proportion of white people",
+                           "The proportion of black people",
+                           "Number of voters supporting Republican by 1000")) %>% 
+  rename(`Variable` = variable,
+         `Encoded Name` = coded_name,
+         `Description` = description) %>% 
+  kable(booktabs = TRUE) %>% 
+  column_spec(column = 3, width = "6cm")
+```
 
@@
 In addition, the data also contains the voting pattern in each county.
-The @tbl-summary_variables summarizes all the important variables in this paper.
-
-```{r}
-#| echo: false
-#| message: false
-#| label: tbl-summary_variables
-#| tbl-cap: Summary of important variables in the model
-data.frame(variable = c("High education attainment",
-                        "Income Percentile",
-                        "People with no insurance",
-                        "People with private insurance",
-                        "Males",
-                        "People aged above 85",
-                        "White people",
-                        "Black people",
-                        "People supporting Republican"),
-           coded_name = c("high_education",
-                          "IncomePctile",
-                          "no_insurance",
-                          "private_insurance",
-                          "males",
-                          "age_85",
-                          "white",
-                          "black",
-                          "rep_rate"),
-           description = c("The proportion of local residents with at least Bachelor degree",
-                           "The mean household income percentile representing the income level for each county",
-                           "The proportion of local residents without insurance",
-                           "The proportion of local residents with private insurance",
-                           "The proportion of male residents",
-                           "The proportion of people aged above 85",
-                           "The proportion of white people",
-                           "The proportion of black people",
-                           "Number of voters supporting Republican by 1000")) %>% 
-  rename(`Variable` = variable,
-         `Encoded Name` = coded_name,
-         `Description` = description) %>% 
-  kable(booktabs = TRUE) %>% 
-  column_spec(column = 3, width = "6cm")
-```
 
@@
 In total, this approach will yield ten models, including the two best-fit models.
-To decide which model is the best, I will evaluate them using testing data to see which one has a lower testing $RMSE$.
+To decide which model is the best, I will perform the 10-fold Cross-Validation to
 
@@
 
-## Expected MSE and Bias-Variance Trade-off
+## 10-fold Cross-Validation
 
-The rationale for partitioning the data into training and testing sets is to validate our model.
-That said, by training a model using training data and then evaluating it using test data, we can determine which model produces the lowest expected MSE and better performance.
+The way to compare the two models is to calculate the error by performing the 10-fold Cross-Validation.
+10-fold Cross-Validation means to to randomly split the data to 10 roughly equal-sized folds and the first one is treated as the validation set.
+Then fit the model on the remaining folds and compute the MSE in the first fold.
+Repeat the process for ten times and the CV error is computed as the averaging of these values.
+That is:
 
 $$
-\mathbb{E}\left[(Y-\hat{f}(x))^2 \mid X=x\right]=\operatorname{Var}(\hat{f}(x))+(\mathbb{E}[\hat{f}(x)]-f(x))^2+\sigma^2
+CV_{(10)}=\frac{1}{10}\sum_{i=1}^{10}MSE_i
 $$
-The above equation shows the expected MSE, which underscores the bias-variance trade-off.
-It tells us the trade-off between the model's complexity and the interpretability.
-An intricate model may increase the bias, but its variance will be higher if we vary the data, and vice-versa.
-
-Hence, by evaluating our models using the testing data, we can know whether we should fit a relatively simple or complicated model.
+Hence, we can compare the CV error for the two models and the model will a lower CV error will perform better.
 
@@
 
-## Comparing the testing error
+## 10 Fold Cross-Validation Error
 
-The above two models are fitted using the training data.
-Comparing the training RMSE from @tbl-all_models1 and @tbl-all_models2, we can find that the model containing the political preference has a lower training RMSE.
-As introduced in the Methods part, to test the prediction performance of the models, I will calculate the test error for the two models to see which is better.
+@tbl-cv_error shows the 10-fold Cross-Validation error of the above two models.
+We can see that the model containing the political preference has a smaller error.
+Therefore, it has a better performance than another one and we should consider the county's political preference when we predict the COVID mortality rate of each county.
 
@@
 #| message: false
-#| label: tbl-test_error
+#| label: tbl-cv_error
 #| tbl-cap: The testing error of the above two models
 
-calculate_testing_error <- function(model, testing_data, outcome) {
-  # Predict the outcomes on the testing data
-  predictions <- predict(model, newdata = testing_data)
-
-  # Compute the residuals
-  residuals <- predictions - testing_data[[outcome]]
-
-  # Compute the MSE
-  mse <- mean(residuals^2)
-
-  return(mse)
-}
+set.seed(0)
+glm.fit <- glm(best1, data =training_data)
+cv_error_1 <- cv.glm(training_data, glm.fit, K=10)$delta[1]
+glm.fit <- glm(best2, data = training_data)
+cv_error_2 <- cv.glm(training_data, glm.fit, K=10)$delta[1]
 
 data.frame(model = c("Model 1", "Model 2"),
-           error = c(calculate_testing_error(best1, testing_data, "death_rate"),
-                     calculate_testing_error(best2, testing_data, "death_rate"))) %>% 
+           error = c(cv_error_1,
+                     cv_error_2)) %>% 
   rename(`Model` = model,
-         `Testinf Error` = error) %>% 
+         `CV Error` = error) %>% 
   kable(booktabs = TRUE)
 ```
-
-@tbl-test_error shows the testing error of the above two models.
-We can observe that the errors for the two models are 1.769733 and 1.675969.
-Therefore, the model containing the political preferences has a lower testing MSE; it has better predictions than the another one.
 

### 2023-12-19 outputs/paper/paper.qmd
Update the final paper

@@
 ---
-title: "Accessing the COVID-19 Mortality Rate in the US: While It Varies Across Different Income Levels, the Impact of Political Preference is Unignorable"
+title: "Examining the COVID-19 Mortality Rate in the US: While It Varies Across Different Income Levels, the Impact of Political Preference is Unignorable"
 title-block-banner: true
@@
 thanks: "Code and data from this analysis are available at: https://github.com/yiliuc/COVID_US_county.git"
-date: "today"
+date: "2023-11-01"
 date-format: "long"
